### 2023-03-13 outputs/paper/paper.qmd
data cleaning 0313

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width=6.5, fig.height=2.8) 
# loading in relevant libraries
library(tidyverse)
library(dplyr, warn.conflicts = FALSE)
library(tidyr)
library(readxl)
library(data.table)
library(lubridate)
library(haven)
library(readxl)
library(ggplot2)
### Input data ###
income_race<-read.csv(here::here("outputs/data/race_income_data.csv"))

ggplot(subset_data_freq, aes(x = income, y = freq, fill = race)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_x_discrete(labels = c("LT $1000","", "$1000 TO 2999","", "$3000 TO 3999","",
                              "$4000 TO 4999","", "$5000 TO 5999","","$6000 TO 6999", "",
                              "$7000 TO 7999","","$8000 TO 9999","","$10000 - 14999","",
                              "$15000 - 19999","","$20000 - 24999","","$25000 OR MORE")) +
  labs(y = "Numbers of people", title = "Numbers of people of Income by race") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))

```

```{r}

```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.

# Data {#sec-data}

Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References

### 2023-03-13 education.R

@@ -1,0 +1,35 @@
+#### Preamble ####
+# Purpose: data cleaning for race
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+# Input data
+raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
+
+# extract the income and race columns
+income_eduyear_data <- raw_data_2018[, c("edu_year", "income")]
+
+# calculate the number of income based on race
+income_eduyear_data <- income_eduyear_data %>%
+  group_by(edu_year, income) %>%
+  summarize(freq = n())
+
+# save the data frame as a csv file
+write.csv(income_eduyear_data, file = here::here("outputs/data/income_eduyear_data.csv"), row.names = TRUE)
+
+

### 2023-03-13 scripts/major.R

@@ -1,0 +1,20 @@
+#### Preamble ####
+# Purpose: data cleaning for race
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: myra.li@mail.utoronto.ca/yx.fan@mail.utoronto.ca
+
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+
+# Input data
+raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))

### 2023-03-13 scripts/preparation.R

@@ -1,0 +1,91 @@
+#### Preamble ####
+# Purpose: Preparation for further data cleaning.
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: myra.li@mail.utoronto.ca/yx.fan@mail.utoronto.ca
+
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+
+# Input raw data
+raw_data_2014<-read_dta(here::here("inputs/data/GSS2014.dta"))
+raw_data_2016<-read_dta(here::here("inputs/data/GSS2016.dta"))
+raw_data_2018<-read_dta(here::here("inputs/data/GSS2018.dta"))
+
+# Select only certain columns
+data_2014 <-
+  raw_data_2014 |>
+  select(
+    id,
+    educ,
+    majorcol,
+    rincome,
+    race,
+    sex
+  )
+
+data_2016 <-
+  raw_data_2016 |>
+  select(
+    id,
+    educ,
+    majorcol,
+    rincome,
+    race,
+    sex
+  )
+
+data_2018 <-
+  raw_data_2018 |>
+  select(
+    id,
+    educ,
+    majorcol,
+    rincome,
+    race,
+    sex
+  )
+
+# Delete the line if the value is NA
+data_2014 <- na.omit(data_2014)
+data_2016 <- na.omit(data_2016)
+data_2018 <- na.omit(data_2018)
+
+# Rename the column
+data_2014 <-
+  data_2014 |>
+  rename(
+    edu_year = educ,
+    major = majorcol,
+    income = rincome
+  )
+
+data_2016 <-
+  data_2016 |>
+  rename(
+    edu_year = educ,
+    major = majorcol,
+    income = rincome
+  )
+
+data_2018 <-
+  data_2018 |>
+  rename(
+    edu_year = educ,
+    major = majorcol,
+    income = rincome
+  )
+
+# Save data
+write.csv(data_2014, file = here::here("outputs/data/preparation/GSS2014.csv"), row.names = FALSE)
+write.csv(data_2016, file = here::here("outputs/data/preparation/GSS2016.csv"), row.names = FALSE)
+write.csv(data_2018, file = here::here("outputs/data/preparation/GSS2018.csv"), row.names = FALSE)

### 2023-03-13 scripts/preview.R

@@ -1,61 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-# Loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-
-# Input raw data
-raw_data_2016<-read_dta(here::here("inputs/data/GSS2016.dta"))
-raw_data_2018<-read_dta(here::here("inputs/data/GSS2018.dta"))
-raw_data_2021<-read_dta(here::here("inputs/data/GSS2021.dta"))
-
-### Clean data ###
-data_2016 <-
-  raw_data_2016 |>
-  # Select only certain columns
-  select(
-    id,
-    educ,
-    majorcol,
-    rincome,
-    race,
-    sex
-  )
-data_2018 <-
-  raw_data_2018 |>
-  # Select only certain columns
-  select(
-    id,
-    educ,
-    majorcol,
-    rincome,
-    race,
-    sex
-  )
-data_2021 <-
-  raw_data_2021 |>
-  # Select only certain columns
-  select(
-    id,
-    educ,
-    race,
-    sex
-  )
-write.csv(data_2016, file = here::here("outputs/data/GSS2016.csv"), row.names = FALSE)
-write.csv(data_2018, file = here::here("outputs/data/GSS2018.csv"), row.names = FALSE)
-write.csv(data_2021, file = here::here("outputs/data/GSS2021.csv"), row.names = FALSE)

### 2023-03-13 scripts/race.R

@@ -1,0 +1,35 @@
+#### Preamble ####
+# Purpose: data cleaning for race
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+# Input data
+raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
+
+# extract the income and race columns
+income_race_data <- raw_data_2018[, c("race", "income")]
+
+# calculate the number of income based on race
+income_race_data <- income_race_data %>%
+  group_by(race, income) %>%
+  summarize(freq = n())
+
+# save the data frame as a csv file
+write.csv(subset_data_freq, file = here::here("outputs/data/race_income.csv"), row.names = TRUE)
+
+

### 2023-03-13 outputs/paper/paper.qmd
data cleaning for sex

@@
 ```{r}
+knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(warning = FALSE)
+knitr::opts_chunk$set(message = FALSE)
+knitr::opts_chunk$set(fig.width=6.5, fig.height=2.8) 
+# loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+### Input data ###
+income_race<-read.csv(here::here("outputs/data/race_income_data.csv"))
+
+ggplot(subset_data_freq, aes(x = income, y = freq, fill = edu_year)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_x_discrete(labels = c("LT $1000","", "$1000 TO 2999","", "$3000 TO 3999","",
+                              "$4000 TO 4999","", "$5000 TO 5999","","$6000 TO 6999", "",
+                              "$7000 TO 7999","","$8000 TO 9999","","$10000 - 14999","",
+                              "$15000 - 19999","","$20000 - 24999","","$25000 OR MORE")) +
+  labs(y = "Numbers of people", title = "Numbers of people of Income by race") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
 

### 2023-03-13 education.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: data cleaning for race
+# Purpose: data cleaning for education
 # Author: Rae Zhang, Faustine Fan, Myra Li
 # Data: 11 March 2023 
 # Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca

### 2023-03-13 scripts/race.R

@@ -30,6 +30,6 @@ income_race_data <- income_race_data %>%
   summarize(freq = n())
 
 # save the data frame as a csv file
-write.csv(subset_data_freq, file = here::here("outputs/data/race_income.csv"), row.names = TRUE)
+write.csv(income_race_data, file = here::here("outputs/data/race_income.csv"), row.names = TRUE)
 
 

### 2023-03-13 scripts/sex.R

@@ -1,26 +1,33 @@
 #### Preamble ####
-# Purpose: Cleans.... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: data cleaning for sex
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
 
 
-#### Workspace setup ####
+
+# Loading in relevant libraries
 library(tidyverse)
-# [...UPDATE THIS...]
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
 
-#### Clean data ####
-# [...UPDATE THIS...]
-raw_data <- read_csv("inputs/data/raw_data.csv")
+# Input data
+raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
 
+# extract the income and race columns
+income_sex_data <- raw_data_2018[, c("sex", "income")]
 
+# calculate the number of income based on race
+income_sex_data <- income_sex_data %>%
+  group_by(sex, income) %>%
+  summarize(freq = n())
 
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change cleaned_data to whatever name you end up with at the end of cleaning
-write_csv(cleaned_data, "outputs/data/cleaned_data.csv")
-
+# save the data frame as a csv file
+write.csv(income_sex_data, file = here::here("outputs/data/sex_income.csv"), row.names = TRUE)

### 2023-03-14 scripts/employee_status.R
data cleaning

@@ -22,14 +22,14 @@ library(reshape2)
 raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
 
 # extract the income and race columns
-income_eduyear_data <- raw_data_2018[, c("edu_year", "income")]
+income_employee_status_data <- raw_data_2018[, c("employee_status", "income")]
 
 # calculate the number of income based on race
-income_eduyear_data <- income_eduyear_data %>%
-  group_by(edu_year, income) %>%
+income_employee_status_data <- income_employee_status_data %>%
+  group_by(employee_status, income) %>%
   summarize(freq = n())
 
 # save the data frame as a csv file
-write.csv(income_eduyear_data, file = here::here("outputs/data/income_eduyear_data.csv"), row.names = TRUE)
+write.csv(income_employee_status_data, file = here::here("outputs/data/income_employee_status_data.csv"), row.names = TRUE)
 
 

### 2023-03-14 scripts/industry_sector.R

@@ -22,14 +22,14 @@ library(reshape2)
 raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
 
 # extract the income and race columns
-income_eduyear_data <- raw_data_2018[, c("edu_year", "income")]
+income_industry_sector_data <- raw_data_2018[, c("industry_sector", "income")]
 
 # calculate the number of income based on race
-income_eduyear_data <- income_eduyear_data %>%
-  group_by(edu_year, income) %>%
+income_industry_sector_data <- income_industry_sector_data %>%
+  group_by(industry_sector, income) %>%
   summarize(freq = n())
 
 # save the data frame as a csv file
-write.csv(income_eduyear_data, file = here::here("outputs/data/income_eduyear_data.csv"), row.names = TRUE)
+write.csv(income_industry_sector_data, file = here::here("outputs/data/income_industry_sector_data.csv"), row.names = TRUE)
 
 

### 2023-03-14 scripts/major.R

@@ -1,20 +1,0 @@
-#### Preamble ####
-# Purpose: data cleaning for race
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: myra.li@mail.utoronto.ca/yx.fan@mail.utoronto.ca
-
-
-# Loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-
-# Input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))

### 2023-03-14 scripts/preparation.R

@@ -25,23 +25,13 @@ raw_data_2018<-read_dta(here::here("inputs/data/GSS2018.dta"))
 data_2014 <-
   raw_data_2014 |>
   select(
-    id,
-    educ,
-    majorcol,
-    rincome,
-    race,
-    sex
+    rincome
   )
 
 data_2016 <-
   raw_data_2016 |>
   select(
-    id,
-    educ,
-    majorcol,
-    rincome,
-    race,
-    sex
+    rincome
   )
 
 data_2018 <-
@@ -49,10 +39,12 @@ data_2018 <-
   select(
     id,
     educ,
-    majorcol,
+    wrkslf,
+    wrkgovt,
     rincome,
     race,
-    sex
+    sex,
+    colsci
   )
 
 # Delete the line if the value is NA
@@ -61,28 +53,14 @@ data_2016 <- na.omit(data_2016)
 data_2018 <- na.omit(data_2018)
 
 # Rename the column
-data_2014 <-
-  data_2014 |>
-  rename(
-    edu_year = educ,
-    major = majorcol,
-    income = rincome
-  )
-
-data_2016 <-
-  data_2016 |>
-  rename(
-    edu_year = educ,
-    major = majorcol,
-    income = rincome
-  )
-
 data_2018 <-
   data_2018 |>
   rename(
     edu_year = educ,
-    major = majorcol,
-    income = rincome
+    employee_status = wrkslf,
+    income = rincome,
+    industry_sector = wrkgovt,
+    science_course = colsci
   )
 
 # Save data

### 2023-03-14 scripts/science_course.R

@@ -22,14 +22,14 @@ library(reshape2)
 raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
 
 # extract the income and race columns
-income_eduyear_data <- raw_data_2018[, c("edu_year", "income")]
+income_science_course_data <- raw_data_2018[, c("science_course", "income")]
 
 # calculate the number of income based on race
-income_eduyear_data <- income_eduyear_data %>%
-  group_by(edu_year, income) %>%
+income_science_course_data <- income_science_course_data %>%
+  group_by(science_course, income) %>%
   summarize(freq = n())
 
 # save the data frame as a csv file
-write.csv(income_eduyear_data, file = here::here("outputs/data/income_eduyear_data.csv"), row.names = TRUE)
+write.csv(income_science_course_data, file = here::here("outputs/data/income_science_course_data.csv"), row.names = TRUE)
 
 

### 2023-03-14 outputs/paper/paper.qmd
data cleaning and graph generation for sex and race

--- 2023-03-13 outputs/paper/paper.qmd
@@
 ### Input data ###
-income_race<-read.csv(here::here("outputs/data/race_income_data.csv"))
+income_race<-read.csv(here::here("outputs/data/race_income.csv"))
+income_sex<-read.csv(here::here("outputs/data/sex_income.csv"))
+```
 
-ggplot(subset_data_freq, aes(x = income, y = freq, fill = race)) +
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-raceincome
+#| fig-cap: "U.S. Income Distribution by Race"
+
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+income_race$income <- 
+  factor(income_race$income, 
+         levels = income_label)
+
+#generate graph#
+ggplot(income_race, aes(x = income, y = the_number_of_people, fill = race)) +
   geom_bar(stat = "identity", position = position_dodge()) +
-  scale_x_discrete(labels = c("LT $1000","", "$1000 TO 2999","", "$3000 TO 3999","",
-                              "$4000 TO 4999","", "$5000 TO 5999","","$6000 TO 6999", "",
-                              "$7000 TO 7999","","$8000 TO 9999","","$10000 - 14999","",
-                              "$15000 - 19999","","$20000 - 24999","","$25000 OR MORE")) +
-  labs(y = "Numbers of people", title = "Numbers of people of Income by race") +
+  labs(x = "Income level", y = "Numbers of people", title = "U.S. Income Distribution by Race") +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8))
 
@@
 ```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-raceincome
+#| fig-cap: "U.S. Income Distribution by Sex"
+
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+income_sex$income <- 
+  factor(income_sex$income, 
+         levels = income_label)
+
+#generate graph#
+ggplot(income_sex, aes(x = income, y = the_number_of_people, fill = sex)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  labs(x = "Income level", y = "Numbers of people", title = "U.S. Income Distribution by Sex") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8))
 

### 2023-03-14 scripts/education.R

@@ -30,6 +30,6 @@ income_eduyear_data <- income_eduyear_data %>%
   summarize(freq = n())
 
 # save the data frame as a csv file
-write.csv(income_eduyear_data, file = here::here("outputs/data/income_eduyear_data.csv"), row.names = TRUE)
+write.csv(income_eduyear_data, file = here::here("outputs/data/eduyear_income_data.csv"), row.names = TRUE)
 
 

### 2023-03-14 scripts/race.R

@@ -29,6 +29,43 @@ income_race_data <- income_race_data %>%
   group_by(race, income) %>%
   summarize(freq = n())
 
+# add two empty rows for black income level 4, and black income level 6 to keep consistent
+new_row <- data.frame(race = 2, income = 4, freq = 0)
+income_race_data <- dplyr::bind_rows(income_race_data[1:15,], new_row, income_race_data[16:nrow(income_race_data),])
+new_row <- data.frame(race = 2, income = 6, freq = 0)
+income_race_data <- dplyr::bind_rows(income_race_data[1:17,], new_row, income_race_data[18:nrow(income_race_data),])
+
+# replace the race code to White, Black, Others
+race_replacements <- c("White", "Black", "Others")
+income_race_data $race <- ifelse(income_race_data $race == 1, race_replacements[1],
+                                 ifelse(income_race_data $race == 2, race_replacements[2], 
+                                        race_replacements[3]))
+
+# replace the income level code to White, Black, Others
+income_race_data <- income_race_data %>%
+  mutate(income = case_when(
+    income == 1 ~ "LT $1000",
+    income == 2 ~ "$1000 TO 2999",
+    income == 3 ~ "$3000 TO 3999",
+    income == 4 ~ "$4000 TO 4999",
+    income == 5 ~ "$5000 TO 5999",
+    income == 6 ~ "$6000 TO 6999",
+    income == 7 ~ "$7000 TO 7999",
+    income == 8 ~ "$8000 TO 9999",
+    income == 9 ~ "$10000 - 14999",
+    income == 10 ~ "$15000 - 19999",
+    income == 11 ~ "$20000 - 24999",
+    income == 12 ~ "$25000 OR MORE",
+    TRUE ~ as.character(income)
+  ))
+
+# rename the column
+income_race_data <-
+  income_race_data |>
+  rename(
+    the_number_of_people = freq
+  )
+
 # save the data frame as a csv file
 write.csv(income_race_data, file = here::here("outputs/data/race_income.csv"), row.names = TRUE)
 

### 2023-03-14 scripts/sex.R
data cleaning_sex

@@ -29,5 +29,39 @@ income_sex_data <- income_sex_data %>%
   group_by(sex, income) %>%
   summarize(freq = n())
 
+# add one empty rows for male income level 6, and black income level 6 to keep consistent
+new_row <- data.frame(sex = 1, income = 6, freq = 0)
+income_sex_data <- dplyr::bind_rows(income_sex_data[1:5,], new_row, income_sex_data[6:nrow(income_sex_data),])
+
+# replace the sex code to Male and Female 
+sex_replacements <- c("Male", "Female")
+income_sex_data$sex <- ifelse(income_sex_data$sex == 1, sex_replacements[1],
+                                sex_replacements[2])
+
+# replace the income level code to White, Black, Others
+income_sex_data <- income_sex_data %>%
+  mutate(income = case_when(
+    income == 1 ~ "LT $1000",
+    income == 2 ~ "$1000 TO 2999",
+    income == 3 ~ "$3000 TO 3999",
+    income == 4 ~ "$4000 TO 4999",
+    income == 5 ~ "$5000 TO 5999",
+    income == 6 ~ "$6000 TO 6999",
+    income == 7 ~ "$7000 TO 7999",
+    income == 8 ~ "$8000 TO 9999",
+    income == 9 ~ "$10000 - 14999",
+    income == 10 ~ "$15000 - 19999",
+    income == 11 ~ "$20000 - 24999",
+    income == 12 ~ "$25000 OR MORE",
+    TRUE ~ as.character(income)
+  ))
+
+# rename the column
+income_sex_data <-
+  income_sex_data |>
+  rename(
+    the_number_of_people = freq
+  )
+
 # save the data frame as a csv file
 write.csv(income_sex_data, file = here::here("outputs/data/sex_income.csv"), row.names = TRUE)

### 2023-03-14 outputs/paper/paper.qmd
graph generate and data cleaning for college science course, industry sector, employee status, race and sex.

--- 2023-03-13 outputs/paper/paper.qmd
@@
 ### Input data ###
-income_race<-read.csv(here::here("outputs/data/race_income_data.csv"))
-
-ggplot(subset_data_freq, aes(x = income, y = freq, fill = race)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  scale_x_discrete(labels = c("LT $1000","", "$1000 TO 2999","", "$3000 TO 3999","",
-                              "$4000 TO 4999","", "$5000 TO 5999","","$6000 TO 6999", "",
-                              "$7000 TO 7999","","$8000 TO 9999","","$10000 - 14999","",
-                              "$15000 - 19999","","$20000 - 24999","","$25000 OR MORE")) +
-  labs(y = "Numbers of people", title = "Numbers of people of Income by race") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8))
-
-```
-
-```{r}
-
+income_sex_data_percentage<-read.csv(here::here("outputs/data/sex_income.csv"))
+income_race_data_percentage<-read.csv(here::here("outputs/data/race_income.csv"))
+income_employee_status_data_percentage<-read.csv(here::here("outputs/data/income_employee_status_data.csv"))
+income_industry_sector_data_percentage<-read.csv(here::here("outputs/data/income_industry_sector_data.csv"))
+income_science_course_data_percentage<-read.csv(here::here("outputs/data/income_science_course_data.csv"))
 ```
@@
 # Data {#sec-data}
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-raceincome
+#| fig-cap: "U.S. Income Distribution by Sex"
 
-Our data is of penguins (@fig-bills).
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+income_sex_data_percentage$Income <- 
+  factor(income_sex_data_percentage$Income, 
+         levels = income_label)
+
+# Manually set fill colors and labels
+fill_colors <- c("Male" = "#9FC69F", "Female" = "#D6BABA")
+
+ggplot(income_sex_data_percentage, aes(x = Income, 
+                                        y = The_Number_Of_People, 
+                                        fill = Sex)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = fill_colors) +
+  labs(x = "Income Level", y = "Total Number of People",
+       title = "U.S. Income Distribution by Sex") +
+  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = The_Number_Of_People + 10), 
+            position = position_stack(vjust = 0.5), size = 2) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
+  ylim(0, 420)
+
+```
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| message: false
+#| warning: false
 #| echo: false
+#| label: fig-raceincome
+#| fig-cap: "U.S. Income Distribution by Race"
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+income_race_data_percentage$Income <- 
+  factor(income_race_data_percentage$Income, 
+         levels = income_label)
+
+# Manually set fill colors and labels
+fill_colors <- c("White" = "#6B5B95", "Black" = "#ADD8E6", "Others" =  "#F5E5C0")
+fill_labels <- c("White" = "White", "Black" = "Black", "Others" = "Other")
+
+ggplot(income_race_data_percentage, aes(x = Income, 
+                                        y = The_Number_Of_People, 
+                                        fill = Race)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = fill_colors, labels = fill_labels) +
+  labs(x = "Income Level", y = "Total Number of People",
+       title = "U.S. Income Distribution by Race") +
+  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = The_Number_Of_People + 10), 
+            position = position_stack(vjust = 0.5), size = 2) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
+  ylim(0, 420)
+
 ```
 
-Talk more about it.
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-raceincome
+#| fig-cap: "U.S. Income Distribution by Employee Status"
 
-Also bills and their average (@fig-billssssss).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+income_employee_status_data_percentage$Income <- 
+  factor(income_employee_status_data_percentage$Income, 
+         levels = income_label)
+
+# Manually set fill colors and labels
+fill_colors <- c("Self-employed" = "#A38863", "Someone else" = "#ADD8E6")
+
+ggplot(income_employee_status_data_percentage, aes(x = Income, 
+                                        y = The_Number_Of_People, 
+                                        fill = EmployeeStatus)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = fill_colors) +
+  labs(x = "Income Level", y = "Total Number of People",
+       title = "U.S. Income Distribution by Employee Status") +
+  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = The_Number_Of_People + 10), 
+            position = position_stack(vjust = 0.5), size = 2) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
+  ylim(0, 420)
+
+```
 
 ```{r}
-#| label: fig-billssssss
-#| fig-cap: More bills of penguins
+#| message: false
+#| warning: false
 #| echo: false
-#| fig-width: 8
-#| fig-height: 4
+#| label: fig-raceincome
+#| fig-cap: "U.S. Income Distribution by Industry Sector"
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+income_industry_sector_data_percentage$Income <- 
+  factor(income_industry_sector_data_percentage$Income, 
+         levels = income_label)
+
+# Manually set fill colors and labels
+fill_colors <- c("Government" = "#72b7b7", "Private" = "#c84b4b")
+
+ggplot(income_industry_sector_data_percentage, aes(x = Income, 
+                                        y = The_Number_Of_People, 
+                                        fill = IndustrySector)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = fill_colors) +
+  labs(x = "Income Level", y = "Total Number of People",
+       title = "U.S. Income Distribution by Industry Sector") +
+  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = The_Number_Of_People + 10), 
+            position = position_stack(vjust = 0.5), size = 2) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
+  ylim(0, 420)
 ```
 
-Talk way more about it.
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-raceincome
+#| fig-cap: "U.S. Income Distribution by College Level Science Courses"
 
-# Model
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+income_science_course_data_percentage$Income <- 
+  factor(income_science_course_data_percentage$Income, 
+         levels = income_label)
 
-$$
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
-$$ {#eq-bayes}
+# Manually set fill colors and labels
+fill_colors <- c("Yes" = "#DCE3A8", "No" = "#C8BCE5")
 
-@eq-bayes seems useful, eh?
-
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+ggplot(income_science_course_data_percentage, aes(x = Income, 
+                                        y = The_Number_Of_People, 
+                                        fill = ScienceCourse)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = fill_colors) +
+  labs(x = "Income Level", y = "Total Number of People",
+       title = "U.S. Income Distribution by College Level Science Courses") +
+  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = The_Number_Of_People + 10), 
+            position = position_stack(vjust = 0.5), size = 2) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
+  ylim(0, 420)
+```
 

### 2023-03-14 myratrytry.R

@@ -1,0 +1,29 @@
+# create example data frame
+df <- data.frame(
+  race = c("White", "Black", "Other", "White", "Black", "Other"),
+  income = c("low", "low", "low", "high", "high", "high"),
+  the_number_of_people = c(100, 150, 50, 200, 300, 100)
+)
+
+# aggregate the_number_of_people column for each income level
+df_sum <- df %>%
+  group_by(income) %>%
+  summarise(total = sum(the_number_of_people))
+
+# merge with original data frame to get percentage
+df_percentage <- df %>%
+  left_join(df_sum, by = "income") %>%
+  group_by(race, income) %>%
+  mutate(percentage = the_number_of_people / total * 100) %>%
+  select(-total)
+
+# create stacked bar chart with total number of people as y-axis
+ggplot(df_percentage, aes(x = income, y = the_number_of_people, fill = race)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = c("White" = "blue", "Black" = "red", "Other" = "green")) +
+  labs(x = "Income Level", y = "Total Number of People",
+       title = "Percentage of Each Race by Income Level") +
+  geom_text(aes(label = paste0(round(percentage, 1), "%"), y = cumsum(percentage) - percentage / 2), 
+            position = position_stack(vjust = 0.5), size = 3)
+
+df

### 2023-03-14 scripts/education.R

@@ -29,6 +29,26 @@ income_eduyear_data <- income_eduyear_data %>%
   group_by(edu_year, income) %>%
   summarize(freq = n())
 
+# rename the column
+income_eduyear_data <-
+  income_eduyear_data |>
+  rename(
+    The_Number_Of_People = freq,
+    EduYear = edu_year,
+    Income = income
+  )
+
+# aggregate the_number_of_people column for each income level
+The_Number_Of_People_Sum <- income_eduyear_data %>%
+  group_by(Income) %>%
+  summarise(total = sum(The_Number_Of_People))
+
+# merge with original data frame to get percentage
+income_eduyear_data_percentage <- income_eduyear_data %>%
+  left_join(The_Number_Of_People_Sum, by = "Income") %>%
+  group_by(EduYear, Income) %>%
+  mutate(Percentage = The_Number_Of_People / total * 100) %>%
+  select(-total)
 # save the data frame as a csv file
 write.csv(income_eduyear_data, file = here::here("outputs/data/eduyear_income_data.csv"), row.names = TRUE)
 

### 2023-03-14 scripts/employee_status.R

@@ -29,7 +29,52 @@ income_employee_status_data <- income_employee_status_data %>%
   group_by(employee_status, income) %>%
   summarize(freq = n())
 
+# replace the employee status code 
+employee_status_replacements <- c("Self-employed", "Someone else")
+income_employee_status_data$employee_status <- ifelse(income_employee_status_data$employee_status == 1, 
+                                          employee_status_replacements[1],
+                                          employee_status_replacements[2])
+
+# replace the income level code
+income_employee_status_data <- income_employee_status_data %>%
+  mutate(income = case_when(
+    income == 1 ~ "LT $1000",
+    income == 2 ~ "$1000 TO 2999",
+    income == 3 ~ "$3000 TO 3999",
+    income == 4 ~ "$4000 TO 4999",
+    income == 5 ~ "$5000 TO 5999",
+    income == 6 ~ "$6000 TO 6999",
+    income == 7 ~ "$7000 TO 7999",
+    income == 8 ~ "$8000 TO 9999",
+    income == 9 ~ "$10000 - 14999",
+    income == 10 ~ "$15000 - 19999",
+    income == 11 ~ "$20000 - 24999",
+    income == 12 ~ "$25000 OR MORE",
+    TRUE ~ as.character(income)
+  ))
+
+# rename the column
+income_employee_status_data <-
+  income_employee_status_data |>
+  rename(
+    The_Number_Of_People = freq,
+    EmployeeStatus = employee_status,
+    Income = income
+  )
+
+# aggregate the_number_of_people column for each income level
+The_Number_Of_People_Sum <- income_employee_status_data %>%
+  group_by(Income) %>%
+  summarise(total = sum(The_Number_Of_People))
+
+# merge with original data frame to get percentage
+income_employee_status_data_percentage <- income_employee_status_data %>%
+  left_join(The_Number_Of_People_Sum, by = "Income") %>%
+  group_by(EmployeeStatus,Income) %>%
+  mutate(Percentage = The_Number_Of_People / total * 100) %>%
+  select(-total)
+
 # save the data frame as a csv file
-write.csv(income_employee_status_data, file = here::here("outputs/data/income_employee_status_data.csv"), row.names = TRUE)
+write.csv(income_employee_status_data_percentage, file = here::here("outputs/data/income_employee_status_data.csv"), row.names = TRUE)
 
 

### 2023-03-14 scripts/industry_sector.R

@@ -29,7 +29,52 @@ income_industry_sector_data <- income_industry_sector_data %>%
   group_by(industry_sector, income) %>%
   summarize(freq = n())
 
+# replace the industry sector code 
+industry_sector_replacements <- c("Government", "Private")
+income_industry_sector_data$industry_sector <- ifelse(income_industry_sector_data$industry_sector == 1, 
+                                                      industry_sector_replacements[1],
+                                                      industry_sector_replacements[2])
+
+# replace the income level code
+income_industry_sector_data <- income_industry_sector_data %>%
+  mutate(income = case_when(
+    income == 1 ~ "LT $1000",
+    income == 2 ~ "$1000 TO 2999",
+    income == 3 ~ "$3000 TO 3999",
+    income == 4 ~ "$4000 TO 4999",
+    income == 5 ~ "$5000 TO 5999",
+    income == 6 ~ "$6000 TO 6999",
+    income == 7 ~ "$7000 TO 7999",
+    income == 8 ~ "$8000 TO 9999",
+    income == 9 ~ "$10000 - 14999",
+    income == 10 ~ "$15000 - 19999",
+    income == 11 ~ "$20000 - 24999",
+    income == 12 ~ "$25000 OR MORE",
+    TRUE ~ as.character(income)
+  ))
+
+# rename the column
+income_industry_sector_data <-
+  income_industry_sector_data |>
+  rename(
+    The_Number_Of_People = freq,
+    IndustrySector = industry_sector,
+    Income = income
+  )
+
+# aggregate the_number_of_people column for each income level
+The_Number_Of_People_Sum <- income_industry_sector_data %>%
+  group_by(Income) %>%
+  summarise(total = sum(The_Number_Of_People))
+
+# merge with original data frame to get percentage
+income_industry_sector_data_percentage <- income_industry_sector_data %>%
+  left_join(The_Number_Of_People_Sum, by = "Income") %>%
+  group_by(IndustrySector,Income) %>%
+  mutate(Percentage = The_Number_Of_People / total * 100) %>%
+  select(-total)
+
 # save the data frame as a csv file
-write.csv(income_industry_sector_data, file = here::here("outputs/data/income_industry_sector_data.csv"), row.names = TRUE)
+write.csv(income_industry_sector_data_percentage, file = here::here("outputs/data/income_industry_sector_data.csv"), row.names = TRUE)
 
 

### 2023-03-14 scripts/race.R

@@ -41,7 +41,7 @@ income_race_data $race <- ifelse(income_race_data $race == 1, race_replacements[
                                  ifelse(income_race_data $race == 2, race_replacements[2], 
                                         race_replacements[3]))
 
-# replace the income level code to White, Black, Others
+# replace the income level code
 income_race_data <- income_race_data %>%
   mutate(income = case_when(
     income == 1 ~ "LT $1000",
@@ -63,10 +63,24 @@ income_race_data <- income_race_data %>%
 income_race_data <-
   income_race_data |>
   rename(
-    the_number_of_people = freq
+    The_Number_Of_People = freq,
+    Race = race,
+    Income = income
   )
 
+# aggregate the_number_of_people column for each income level
+The_Number_Of_People_Sum <- income_race_data %>%
+  group_by(Income) %>%
+  summarise(total = sum(The_Number_Of_People))
+
+# merge with original data frame to get percentage
+income_race_data_percentage <- income_race_data %>%
+  left_join(The_Number_Of_People_Sum, by = "Income") %>%
+  group_by(Race, Income) %>%
+  mutate(Percentage = The_Number_Of_People / total * 100) %>%
+  select(-total)
+
 # save the data frame as a csv file
-write.csv(income_race_data, file = here::here("outputs/data/race_income.csv"), row.names = TRUE)
+write.csv(income_race_data_percentage, file = here::here("outputs/data/race_income.csv"), row.names = TRUE)
 
 

### 2023-03-14 scripts/science_course.R

@@ -29,7 +29,52 @@ income_science_course_data <- income_science_course_data %>%
   group_by(science_course, income) %>%
   summarize(freq = n())
 
+# replace the science course code 
+income_science_course_data_replacements <- c("Yes", "No")
+income_science_course_data$science_course <- ifelse(income_science_course_data$science_course == 1, 
+                                                     income_science_course_data_replacements[1],
+                                                     income_science_course_data_replacements[2])
+
+# replace the income level code
+income_science_course_data <- income_science_course_data %>%
+  mutate(income = case_when(
+    income == 1 ~ "LT $1000",
+    income == 2 ~ "$1000 TO 2999",
+    income == 3 ~ "$3000 TO 3999",
+    income == 4 ~ "$4000 TO 4999",
+    income == 5 ~ "$5000 TO 5999",
+    income == 6 ~ "$6000 TO 6999",
+    income == 7 ~ "$7000 TO 7999",
+    income == 8 ~ "$8000 TO 9999",
+    income == 9 ~ "$10000 - 14999",
+    income == 10 ~ "$15000 - 19999",
+    income == 11 ~ "$20000 - 24999",
+    income == 12 ~ "$25000 OR MORE",
+    TRUE ~ as.character(income)
+  ))
+
+# rename the column
+income_science_course_data <-
+  income_science_course_data |>
+  rename(
+    The_Number_Of_People = freq,
+    ScienceCourse = science_course,
+    Income = income
+  )
+
+# aggregate the_number_of_people column for each income level
+The_Number_Of_People_Sum <- income_science_course_data %>%
+  group_by(Income) %>%
+  summarise(total = sum(The_Number_Of_People))
+
+# merge with original data frame to get percentage
+income_science_course_data_percentage <- income_science_course_data %>%
+  left_join(The_Number_Of_People_Sum, by = "Income") %>%
+  group_by(ScienceCourse,Income) %>%
+  mutate(Percentage = The_Number_Of_People / total * 100) %>%
+  select(-total)
+
 # save the data frame as a csv file
-write.csv(income_science_course_data, file = here::here("outputs/data/income_science_course_data.csv"), row.names = TRUE)
+write.csv(income_science_course_data_percentage, file = here::here("outputs/data/income_science_course_data.csv"), row.names = TRUE)
 
 

### 2023-03-14 scripts/sex.R

@@ -38,7 +38,7 @@ sex_replacements <- c("Male", "Female")
 income_sex_data$sex <- ifelse(income_sex_data$sex == 1, sex_replacements[1],
                                 sex_replacements[2])
 
-# replace the income level code to White, Black, Others
+# replace the income level code
 income_sex_data <- income_sex_data %>%
   mutate(income = case_when(
     income == 1 ~ "LT $1000",
@@ -60,8 +60,22 @@ income_sex_data <- income_sex_data %>%
 income_sex_data <-
   income_sex_data |>
   rename(
-    the_number_of_people = freq
+    The_Number_Of_People = freq,
+    Sex = sex,
+    Income = income
   )
 
+# aggregate the_number_of_people column for each income level
+The_Number_Of_People_Sum <- income_sex_data %>%
+  group_by(Income) %>%
+  summarise(total = sum(The_Number_Of_People))
+
+# merge with original data frame to get percentage
+income_sex_data_percentage <- income_sex_data %>%
+  left_join(The_Number_Of_People_Sum, by = "Income") %>%
+  group_by(Sex, Income) %>%
+  mutate(Percentage = The_Number_Of_People / total * 100) %>%
+  select(-total)
+
 # save the data frame as a csv file
-write.csv(income_sex_data, file = here::here("outputs/data/sex_income.csv"), row.names = TRUE)
+write.csv(income_sex_data_percentage, file = here::here("outputs/data/sex_income.csv"), row.names = TRUE)

### 2023-03-14 outputs/paper/paper.qmd
data cleaning and generate graph for income, race, sex, industry sector, employee status. Data cleaning for eduyear and income.

@@
 ### Input data ###
+income_percentage<-read.csv(here::here("outputs/data/income_3years.csv"))
 income_sex_data_percentage<-read.csv(here::here("outputs/data/sex_income.csv"))
@@
 # Data {#sec-data}
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-raceincome
+#| fig-cap: "U.S. Income Distribution from 2014 to 2018"
+
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+income_percentage$Income <- 
+  factor(income_percentage$Income, 
+         levels = income_label)
+
+# Manually set fill colors and labels
+fill_colors <- c("2014" = "#6B5B95", "2016" = "#ADD8E6", "2018" =  "#F5E5C0")
+
+ggplot(income_percentage, aes(x = Income, 
+                              y = Number, 
+                              fill = Year)) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = fill_colors) +
+  labs(x = "Income Level", y = "Total Number of People",
+       title = "U.S. Income Distribution by Year") +
+  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = Number + 10), 
+            position = position_stack(vjust = 0.5), size = 2) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
+  ylim(0, 420)
+
+```
+
 ```{r}

### 2023-03-14 myratrytry.R

@@ -1,29 +1,0 @@
-# create example data frame
-df <- data.frame(
-  race = c("White", "Black", "Other", "White", "Black", "Other"),
-  income = c("low", "low", "low", "high", "high", "high"),
-  the_number_of_people = c(100, 150, 50, 200, 300, 100)
-)
-
-# aggregate the_number_of_people column for each income level
-df_sum <- df %>%
-  group_by(income) %>%
-  summarise(total = sum(the_number_of_people))
-
-# merge with original data frame to get percentage
-df_percentage <- df %>%
-  left_join(df_sum, by = "income") %>%
-  group_by(race, income) %>%
-  mutate(percentage = the_number_of_people / total * 100) %>%
-  select(-total)
-
-# create stacked bar chart with total number of people as y-axis
-ggplot(df_percentage, aes(x = income, y = the_number_of_people, fill = race)) +
-  geom_bar(stat = "identity", position = "stack") +
-  scale_fill_manual(values = c("White" = "blue", "Black" = "red", "Other" = "green")) +
-  labs(x = "Income Level", y = "Total Number of People",
-       title = "Percentage of Each Race by Income Level") +
-  geom_text(aes(label = paste0(round(percentage, 1), "%"), y = cumsum(percentage) - percentage / 2), 
-            position = position_stack(vjust = 0.5), size = 3)
-
-df

### 2023-03-14 scripts/00_preparation.R

@@ -2,7 +2,7 @@
 # Purpose: Preparation for further data cleaning.
 # Author: Rae Zhang, Faustine Fan, Myra Li
 # Data: 11 March 2023 
-# Contact: myra.li@mail.utoronto.ca/yx.fan@mail.utoronto.ca
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
 
 
 # Loading in relevant libraries

### 2023-03-14 scripts/education.R

@@ -5,7 +5,6 @@
 # Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
 
 
-
 # Loading in relevant libraries
 library(tidyverse)
 library(dplyr, warn.conflicts = FALSE)
@@ -49,7 +48,8 @@ income_eduyear_data_percentage <- income_eduyear_data %>%
   group_by(EduYear, Income) %>%
   mutate(Percentage = The_Number_Of_People / total * 100) %>%
   select(-total)
+
 # save the data frame as a csv file
-write.csv(income_eduyear_data, file = here::here("outputs/data/eduyear_income_data.csv"), row.names = TRUE)
+write.csv(income_eduyear_data_percentage, file = here::here("outputs/data/eduyear_income_data.csv"), row.names = TRUE)
 
 

### 2023-03-14 scripts/employee_status.R

@@ -1,9 +1,8 @@
 #### Preamble ####
-# Purpose: data cleaning for education
+# Purpose: data cleaning for employee status
 # Author: Rae Zhang, Faustine Fan, Myra Li
 # Data: 11 March 2023 
 # Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
 
 
 # Loading in relevant libraries

### 2023-03-14 scripts/income.R

@@ -1,0 +1,89 @@
+#### Preamble ####
+# Purpose: data cleaning for three year income
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+# Input data
+income_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) %>% select(income)
+income_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) %>% select(rincome)
+income_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) %>% select(rincome)
+
+# rename the column
+income_2016 <-
+  income_2016|>
+  rename(
+    income = rincome
+  )
+income_2014 <-
+  income_2014|>
+  rename(
+    income = rincome
+  )
+
+# Combine the data frames and add a column for the file name
+income_data <- bind_rows(
+  data.frame(year = "2018", income = income_2018),
+  data.frame(year = "2016", rincome = income_2016),
+  data.frame(year = "2014", rincome = income_2014))
+
+# group by year and income, count the number
+income_data_count <- income_data %>%
+  group_by(year, income) %>%
+  count() %>%
+  rename(number = n)
+
+# replace the income level code
+income_data_count <- income_data_count %>%
+  mutate(income = case_when(
+    income == 1 ~ "LT $1000",
+    income == 2 ~ "$1000 TO 2999",
+    income == 3 ~ "$3000 TO 3999",
+    income == 4 ~ "$4000 TO 4999",
+    income == 5 ~ "$5000 TO 5999",
+    income == 6 ~ "$6000 TO 6999",
+    income == 7 ~ "$7000 TO 7999",
+    income == 8 ~ "$8000 TO 9999",
+    income == 9 ~ "$10000 - 14999",
+    income == 10 ~ "$15000 - 19999",
+    income == 11 ~ "$20000 - 24999",
+    income == 12 ~ "$25000 OR MORE",
+    TRUE ~ as.character(income)
+  ))
+
+# rename the column
+income_data_count <-
+  income_data_count|>
+  rename(
[…]

### 2023-03-14 scripts/industry_sector.R

@@ -1,9 +1,8 @@
 #### Preamble ####
-# Purpose: data cleaning for education
+# Purpose: data cleaning for industry sector
 # Author: Rae Zhang, Faustine Fan, Myra Li
 # Data: 11 March 2023 
 # Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
 
 
 # Loading in relevant libraries

### 2023-03-14 scripts/race.R

@@ -5,7 +5,6 @@
 # Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
 
 
-
 # Loading in relevant libraries
 library(tidyverse)
 library(dplyr, warn.conflicts = FALSE)

### 2023-03-14 scripts/science_course.R

@@ -1,9 +1,8 @@
 #### Preamble ####
-# Purpose: data cleaning for education
+# Purpose: data cleaning for college science course
 # Author: Rae Zhang, Faustine Fan, Myra Li
 # Data: 11 March 2023 
 # Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
 
 
 # Loading in relevant libraries

### 2023-03-14 outputs/paper/paper.qmd
graphs and data cleaning, except education year

--- 2023-03-14 outputs/paper/paper.qmd
@@
 knitr::opts_chunk$set(fig.width=6.5, fig.height=2.8)
+
 # loading in relevant libraries
+library(RColorBrewer)
 library(tidyverse)
@@
 library(ggplot2)
+
 ### Input data ###
-income_race<-read.csv(here::here("outputs/data/race_income.csv"))
-income_sex<-read.csv(here::here("outputs/data/sex_income.csv"))
+income_data_percentage<-read.csv(here::here("outputs/data/income_3years.csv"))
+sex_data<-read.csv(here::here("outputs/data/sex_data.csv"))
+race_data<-read.csv(here::here("outputs/data/race_data.csv"))
+employee_status_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
+industry_sector_data<-read.csv(here::here("outputs/data/industry_sector_data.csv"))
+science_course_data<-read.csv(here::here("outputs/data/science_course_data.csv"))
+
 ```
+
+# Introduction
+
+You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
+
+```{r, fig.height=8}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-1
+#| fig-cap: "U.S. Income Distribution from 2014 to 2018"
+
+# Make sure the order of the income level is the order we wanted
+income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
+
+income_data_percentage$Income <- factor(income_data_percentage$Income, levels = income_label)
+
+# Define color palette with 3 colors
+fill_colors <- brewer.pal(n = 3, name = "RdYlBu")
+
+# Generate graph
+ggplot(income_data_percentage, aes(x = Income, y = Number, fill = as.factor(Year))) +
+  geom_bar(stat = "identity", position = "stack") +
+  scale_fill_manual(values = fill_colors) +
+  labs(x = "Income Level", y = "Total Number of People", title = "U.S. Income Distribution by Year") +
+  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = Number + 10), position = position_stack(vjust = 0.5), size = 2) +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
+  ylim(0, 2400)
+
+```
+
+# Data {#sec-data}
 
@@
 #| echo: false
-#| label: fig-raceincome
-#| fig-cap: "U.S. Income Distribution by Race"
+#| label: fig-2
+#| fig-cap: "U.S. Income Distribution by Sex"
 
-# Make sure the order of the income level is the order we wanted
-income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
-income_race$income <- 
-  factor(income_race$income, 
-         levels = income_label)
+# calculate the percentage of each sex group for each year
+sex_data <- sex_data %>% 
+  group_by(year) %>% 
+  mutate(percentage = 100 * people / sum(people))
 
-#generate graph#
-ggplot(income_race, aes(x = income, y = the_number_of_people, fill = race)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  labs(x = "Income level", y = "Numbers of people", title = "U.S. Income Distribution by Race") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8))
+# generate graph
+ggplot(sex_data, aes(x = year, y = people, 
+                                 fill = sex)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 1), size = 3) +
+  labs(x = "Year", y = "People", fill = "Sex",title = "sex")
 
@@
 #| echo: false
-#| label: fig-raceincome
-#| fig-cap: "U.S. Income Distribution by Sex"
+#| label: fig-3
+#| fig-cap: "U.S. Income Distribution by Race"
 
-# Make sure the order of the income level is the order we wanted
-income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
-income_sex$income <- 
-  factor(income_sex$income, 
-         levels = income_label)
+# calculate the percentage of each race group for each year
+race_data <- race_data %>% 
+  group_by(year) %>% 
+  mutate(percentage = 100 * people / sum(people))
 
-#generate graph#
-ggplot(income_sex, aes(x = income, y = the_number_of_people, fill = sex)) +
-  geom_bar(stat = "identity", position = position_dodge()) +
-  labs(x = "Income level", y = "Numbers of people", title = "U.S. Income Distribution by Sex") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8))
+# generate graph
+ggplot(race_data, aes(x = year, y = people, 
+                                 fill = race)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 2), size = 3) +
+  labs(x = "Year", y = "People", fill = "Race",title = "race")
 
@@
 
-# Introduction
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-4
+#| fig-cap: "U.S. Income Distribution by Employee Status"
 
-You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
+# calculate the percentage of each employee status group for each year
+employee_status_data <- employee_status_data %>% 
+  group_by(year) %>% 
+  mutate(percentage = 100 * people / sum(people))
 
-# Data {#sec-data}
+# generate graph
+ggplot(employee_status_data, aes(x = year, y = people, 
+                                 fill = employee_status)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
+  labs(x = "Year", y = "People", fill = "Employee Status",title = "employee status")
 
-Our data is of penguins (@fig-bills).
+```
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| message: false
+#| warning: false
 #| echo: false
+#| label: fig-5
+#| fig-cap: "U.S. Income Distribution by Industry Sector"
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+# calculate the percentage of each industry sector group for each year
+industry_sector_data <- industry_sector_data %>% 
+  group_by(year) %>% 
+  mutate(percentage = 100 * people / sum(people))
+
+# generate graph
+ggplot(industry_sector_data, aes(x = year, y = people, 
+                                 fill = industry_sector)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
+  labs(x = "Year", y = "People", fill = "Industry Sector",title = "industry sector")
+
 ```
 
-Talk more about it.
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-6
+#| fig-cap: "U.S. Income Distribution by College Level Science Courses"
 
-Also bills and their average (@fig-billssssss).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)
+# calculate the percentage of taking college level science courses for each year
+science_course_data <- science_course_data %>% 
+  group_by(year) %>% 
+  mutate(percentage = 100 * people / sum(people))
 
-```{r}
-#| label: fig-billssssss
-#| fig-cap: More bills of penguins
-#| echo: false
-#| fig-width: 8
-#| fig-height: 4
+# generate graph
+ggplot(science_course_data, aes(x = year, y = people, 
+                                 fill = science_course)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
+  labs(x = "Year", y = "People", fill = "Whether taking science course",title = "Science course")
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
 ```
-
-Talk way more about it.
-
-# Model
-
-$$
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
-$$ {#eq-bayes}
-
-@eq-bayes seems useful, eh?
-
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
 

### 2023-03-14 scripts/00_preparation.R

@@ -25,13 +25,27 @@ raw_data_2018<-read_dta(here::here("inputs/data/GSS2018.dta"))
 data_2014 <-
   raw_data_2014 |>
   select(
-    rincome
+    id,
+    educ,
+    wrkslf,
+    wrkgovt,
+    rincome,
+    race,
+    sex,
+    colsci
   )
 
 data_2016 <-
   raw_data_2016 |>
   select(
-    rincome
+    id,
+    educ,
+    wrkslf,
+    wrkgovt,
+    rincome,
+    race,
+    sex,
+    colsci
   )
 
 data_2018 <-
@@ -53,6 +67,26 @@ data_2016 <- na.omit(data_2016)
 data_2018 <- na.omit(data_2018)
 
 # Rename the column
+data_2016 <-
+  data_2016|>
+  rename(
+    edu_year = educ,
+    employee_status = wrkslf,
+    income = rincome,
+    industry_sector = wrkgovt,
+    science_course = colsci
+  )
+
+data_2014 <-
+  data_2014|>
+  rename(
+    edu_year = educ,
+    employee_status = wrkslf,
+    income = rincome,
+    industry_sector = wrkgovt,
+    science_course = colsci
+  )
+
 data_2018 <-
   data_2018 |>
   rename(

### 2023-03-14 scripts/income.R

@@ -19,26 +19,14 @@ library(reshape2)
 
 # Input data
 income_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) %>% select(income)
-income_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) %>% select(rincome)
-income_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) %>% select(rincome)
-
-# rename the column
-income_2016 <-
-  income_2016|>
-  rename(
-    income = rincome
-  )
-income_2014 <-
-  income_2014|>
-  rename(
-    income = rincome
-  )
+income_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) %>% select(income)
+income_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) %>% select(income)
 
 # Combine the data frames and add a column for the file name
 income_data <- bind_rows(
   data.frame(year = "2018", income = income_2018),
-  data.frame(year = "2016", rincome = income_2016),
-  data.frame(year = "2014", rincome = income_2014))
+  data.frame(year = "2016", income = income_2016),
+  data.frame(year = "2014", income = income_2014))
 
 # group by year and income, count the number
 income_data_count <- income_data %>%
@@ -75,13 +63,13 @@ income_data_count <-
 
 # aggregate number column for each income level
 Number_Sum <- income_data_count %>%
-  group_by(Income) %>%
+  group_by(Year) %>%
   summarise(total = sum(Number))
 
 # merge with original data frame to get percentage
 income_data_percentage <- income_data_count %>%
-  left_join(Number_Sum, by = "Income") %>%
-  group_by(Year, Income) %>%
+  left_join(Number_Sum, by = "Year") %>%
+  group_by(Income, Year) %>%
   mutate(Percentage = Number/ total * 100) %>%
   select(-total)
 

### 2023-03-14 scripts/industry_sector.R

@@ -4,8 +4,7 @@
 # Data: 11 March 2023 
 # Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
 
-
-# Loading in relevant libraries
+# loading in relevant libraries
 library(tidyverse)
 library(dplyr, warn.conflicts = FALSE)
 library(tidyr)
@@ -17,63 +16,37 @@ library(readxl)
 library(ggplot2)
 library(reshape2)
 
-# Input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
+# input data
+raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
+raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
+raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
 
-# extract the income and race columns
-income_industry_sector_data <- raw_data_2018[, c("industry_sector", "income")]
+# extract the income and industry sector columns
+industry_sector_2018 <- raw_data_2018[, c("industry_sector", "income")]
+industry_sector_2016 <- raw_data_2016[, c("industry_sector", "income")]
+industry_sector_2014 <- raw_data_2014[, c("industry_sector", "income")]
 
-# calculate the number of income based on race
-income_industry_sector_data <- income_industry_sector_data %>%
-  group_by(industry_sector, income) %>%
-  summarize(freq = n())
+# combine three files into one
+industry_sector_data <- bind_rows(
+  industry_sector_2018 %>% mutate(year = "2018"),
+  industry_sector_2016 %>% mutate(year = "2016"),
+  industry_sector_2014 %>% mutate(year = "2014")
+)
+
+# only keep the income level 12
+# calcualte the number of each industry sector by year
+industry_sector_data <- industry_sector_data %>% 
+  filter(income == 12) %>% 
+  group_by(industry_sector, year) %>% 
+  summarize(people = n())
 
 # replace the industry sector code 
 industry_sector_replacements <- c("Government", "Private")
-income_industry_sector_data$industry_sector <- ifelse(income_industry_sector_data$industry_sector == 1, 
-                                                      industry_sector_replacements[1],
-                                                      industry_sector_replacements[2])
-
-# replace the income level code
[…]

### 2023-03-14 scripts/race.R

@@ -4,82 +4,37 @@
 # Data: 11 March 2023 
 # Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
 
-
-# Loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-library(reshape2)
-
-# Input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
+# input data
+raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
+raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
+raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
 
 # extract the income and race columns
-income_race_data <- raw_data_2018[, c("race", "income")]
+race_2018 <- raw_data_2018[, c("race", "income")]
+race_2016 <- raw_data_2016[, c("race", "income")]
+race_2014 <- raw_data_2014[, c("race", "income")]
 
-# calculate the number of income based on race
-income_race_data <- income_race_data %>%
-  group_by(race, income) %>%
-  summarize(freq = n())
+# combine three files into one
+race_data <- bind_rows(
+  race_2018 %>% mutate(year = "2018"),
+  race_2016 %>% mutate(year = "2016"),
+  race_2014 %>% mutate(year = "2014")
+)
 
-# add two empty rows for black income level 4, and black income level 6 to keep consistent
-new_row <- data.frame(race = 2, income = 4, freq = 0)
-income_race_data <- dplyr::bind_rows(income_race_data[1:15,], new_row, income_race_data[16:nrow(income_race_data),])
-new_row <- data.frame(race = 2, income = 6, freq = 0)
-income_race_data <- dplyr::bind_rows(income_race_data[1:17,], new_row, income_race_data[18:nrow(income_race_data),])
+# only keep the income level 12
+# calcualte the number of each race by year
+race_data <- race_data %>% 
+  filter(income == 12) %>% 
+  group_by(race, year) %>% 
+  summarize(people = n())
 
-# replace the race code to White, Black, Others
-race_replacements <- c("White", "Black", "Others")
-income_race_data $race <- ifelse(income_race_data $race == 1, race_replacements[1],
[…]

### 2023-03-14 scripts/science_course.R

@@ -17,63 +17,37 @@ library(readxl)
 library(ggplot2)
 library(reshape2)
 
-# Input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
+# input data
+raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
+raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
+raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
 
-# extract the income and race columns
-income_science_course_data <- raw_data_2018[, c("science_course", "income")]
+# extract the income and science course columns
+science_course_2018 <- raw_data_2018[, c("science_course", "income")]
+science_course_2016 <- raw_data_2016[, c("science_course", "income")]
+science_course_2014 <- raw_data_2014[, c("science_course", "income")]
 
-# calculate the number of income based on race
-income_science_course_data <- income_science_course_data %>%
-  group_by(science_course, income) %>%
-  summarize(freq = n())
+# combine three files into one
+science_course_data <- bind_rows(
+  science_course_2018 %>% mutate(year = "2018"),
+  science_course_2016 %>% mutate(year = "2016"),
+  science_course_2014 %>% mutate(year = "2014")
+)
 
-# replace the science course code 
-income_science_course_data_replacements <- c("Yes", "No")
-income_science_course_data$science_course <- ifelse(income_science_course_data$science_course == 1, 
-                                                     income_science_course_data_replacements[1],
-                                                     income_science_course_data_replacements[2])
+# only keep the income level 12
+# calcualte the number of college science course by year
+science_course_data <- science_course_data %>% 
+  filter(income == 12) %>% 
+  group_by(science_course, year) %>% 
+  summarize(people = n())
 
-# replace the income level code
-income_science_course_data <- income_science_course_data %>%
-  mutate(income = case_when(
-    income == 1 ~ "LT $1000",
-    income == 2 ~ "$1000 TO 2999",
-    income == 3 ~ "$3000 TO 3999",
-    income == 4 ~ "$4000 TO 4999",
-    income == 5 ~ "$5000 TO 5999",
-    income == 6 ~ "$6000 TO 6999",
-    income == 7 ~ "$7000 TO 7999",
-    income == 8 ~ "$8000 TO 9999",
-    income == 9 ~ "$10000 - 14999",
-    income == 10 ~ "$15000 - 19999",
[…]

### 2023-03-14 scripts/sex.R

@@ -17,64 +17,35 @@ library(readxl)
 library(ggplot2)
 library(reshape2)
 
-# Input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
+# input data
+raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
+raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
+raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
 
-# extract the income and race columns
-income_sex_data <- raw_data_2018[, c("sex", "income")]
+# extract the income and sex columns
+sex_2018 <- raw_data_2018[, c("sex", "income")]
+sex_2016 <- raw_data_2016[, c("sex", "income")]
+sex_2014 <- raw_data_2014[, c("sex", "income")]
 
-# calculate the number of income based on race
-income_sex_data <- income_sex_data %>%
-  group_by(sex, income) %>%
-  summarize(freq = n())
+# combine three files into one
+sex_data <- bind_rows(
+  sex_2018 %>% mutate(year = "2018"),
+  sex_2016 %>% mutate(year = "2016"),
+  sex_2014 %>% mutate(year = "2014")
+)
 
-# add one empty rows for male income level 6, and black income level 6 to keep consistent
-new_row <- data.frame(sex = 1, income = 6, freq = 0)
-income_sex_data <- dplyr::bind_rows(income_sex_data[1:5,], new_row, income_sex_data[6:nrow(income_sex_data),])
+# only keep the income level 12
+# calcualte the number of each sex by year
+sex_data <- sex_data %>% 
+  filter(income == 12) %>% 
+  group_by(sex, year) %>% 
+  summarize(people = n())
 
-# replace the sex code to Male and Female 
+# replace the sex code 
 sex_replacements <- c("Male", "Female")
-income_sex_data$sex <- ifelse(income_sex_data$sex == 1, sex_replacements[1],
-                                sex_replacements[2])
-
-# replace the income level code
-income_sex_data <- income_sex_data %>%
-  mutate(income = case_when(
-    income == 1 ~ "LT $1000",
-    income == 2 ~ "$1000 TO 2999",
-    income == 3 ~ "$3000 TO 3999",
-    income == 4 ~ "$4000 TO 4999",
-    income == 5 ~ "$5000 TO 5999",
[…]

### 2023-03-14 scripts/test.R

@@ -1,10 +1,3 @@
-#### Preamble ####
-# Purpose: data cleaning for employee status
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-# loading in relevant libraries
 library(tidyverse)
 library(dplyr, warn.conflicts = FALSE)
 library(tidyr)
@@ -16,12 +9,12 @@ library(readxl)
 library(ggplot2)
 library(reshape2)
 
-# input data
+# Input data
 raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
 raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
 raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
 
-# extract the income and employee status columns
+# extract the income and race columns
 employee_status_2018 <- raw_data_2018[, c("employee_status", "income")]
 employee_status_2016 <- raw_data_2016[, c("employee_status", "income")]
 employee_status_2014 <- raw_data_2014[, c("employee_status", "income")]
@@ -38,15 +31,10 @@ employee_status_data <- bind_rows(
 employee_status_data <- income_data %>% 
   filter(income == 12) %>% 
   group_by(employee_status, year) %>% 
-  summarize(people = n())
+  summarize(freq = n())
 
 # replace the employee status code 
 employee_status_replacements <- c("Self-employed", "Someone else")
 employee_status_data$employee_status <- ifelse(employee_status_data$employee_status == 1, 
-                                               employee_status_replacements[1],
-                                               employee_status_replacements[2])
-
-# save the data frame as a csv file
-write.csv(employee_status_data, file = here::here("outputs/data/employee_status_data.csv"), row.names = TRUE)
-
-
+                                                      employee_status_replacements[1],
+                                                      employee_status_replacements[2])

### 2023-03-15 scripts/00_preparation.R
delete unnecessary files

@@ -1,103 +1,0 @@
-#### Preamble ####
-# Purpose: Preparation for further data cleaning.
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-
-# Loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-
-# Input raw data
-raw_data_2014<-read_dta(here::here("inputs/data/GSS2014.dta"))
-raw_data_2016<-read_dta(here::here("inputs/data/GSS2016.dta"))
-raw_data_2018<-read_dta(here::here("inputs/data/GSS2018.dta"))
-
-# Select only certain columns
-data_2014 <-
-  raw_data_2014 |>
-  select(
-    id,
-    educ,
-    wrkslf,
-    wrkgovt,
-    rincome,
-    race,
-    sex,
-    colsci
-  )
-
-data_2016 <-
-  raw_data_2016 |>
-  select(
-    id,
-    educ,
-    wrkslf,
-    wrkgovt,
-    rincome,
-    race,
-    sex,
-    colsci
-  )
-
-data_2018 <-
-  raw_data_2018 |>
-  select(
-    id,
-    educ,
-    wrkslf,
-    wrkgovt,
-    rincome,
-    race,
-    sex,
-    colsci
-  )
-
-# Delete the line if the value is NA
-data_2014 <- na.omit(data_2014)
-data_2016 <- na.omit(data_2016)
-data_2018 <- na.omit(data_2018)
-
-# Rename the column
-data_2016 <-
-  data_2016|>
-  rename(
-    edu_year = educ,
-    employee_status = wrkslf,
-    income = rincome,
-    industry_sector = wrkgovt,
-    science_course = colsci
-  )
-
-data_2014 <-
-  data_2014|>
-  rename(
-    edu_year = educ,
-    employee_status = wrkslf,
-    income = rincome,
-    industry_sector = wrkgovt,
-    science_course = colsci
-  )
-
-data_2018 <-
-  data_2018 |>
-  rename(
-    edu_year = educ,
-    employee_status = wrkslf,
-    income = rincome,
-    industry_sector = wrkgovt,
[…]

### 2023-03-15 scripts/education.R

@@ -1,55 +1,0 @@
-#### Preamble ####
-# Purpose: data cleaning for education
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-
-# Loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-library(reshape2)
-
-# Input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv"))
-
-# extract the income and race columns
-income_eduyear_data <- raw_data_2018[, c("edu_year", "income")]
-
-# calculate the number of income based on race
-income_eduyear_data <- income_eduyear_data %>%
-  group_by(edu_year, income) %>%
-  summarize(freq = n())
-
-# rename the column
-income_eduyear_data <-
-  income_eduyear_data |>
-  rename(
-    The_Number_Of_People = freq,
-    EduYear = edu_year,
-    Income = income
-  )
-
-# aggregate the_number_of_people column for each income level
-The_Number_Of_People_Sum <- income_eduyear_data %>%
-  group_by(Income) %>%
-  summarise(total = sum(The_Number_Of_People))
-
-# merge with original data frame to get percentage
-income_eduyear_data_percentage <- income_eduyear_data %>%
-  left_join(The_Number_Of_People_Sum, by = "Income") %>%
-  group_by(EduYear, Income) %>%
-  mutate(Percentage = The_Number_Of_People / total * 100) %>%
-  select(-total)
-
-# save the data frame as a csv file
-write.csv(income_eduyear_data_percentage, file = here::here("outputs/data/eduyear_income_data.csv"), row.names = TRUE)
-
-

### 2023-03-15 scripts/employee_status.R

@@ -1,52 +1,0 @@
-#### Preamble ####
-# Purpose: data cleaning for employee status
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-# loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-library(reshape2)
-
-# input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
-raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
-raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
-
-# extract the income and employee status columns
-employee_status_2018 <- raw_data_2018[, c("employee_status", "income")]
-employee_status_2016 <- raw_data_2016[, c("employee_status", "income")]
-employee_status_2014 <- raw_data_2014[, c("employee_status", "income")]
-
-# combine three files into one
-employee_status_data <- bind_rows(
-  employee_status_2018 %>% mutate(year = "2018"),
-  employee_status_2016 %>% mutate(year = "2016"),
-  employee_status_2014 %>% mutate(year = "2014")
-)
-
-# only keep the income level 12
-# calcualte the number of each emplyee status by year
-employee_status_data <- income_data %>% 
-  filter(income == 12) %>% 
-  group_by(employee_status, year) %>% 
-  summarize(people = n())
-
-# replace the employee status code 
-employee_status_replacements <- c("Self-employed", "Someone else")
-employee_status_data$employee_status <- ifelse(employee_status_data$employee_status == 1, 
-                                               employee_status_replacements[1],
-                                               employee_status_replacements[2])
-
-# save the data frame as a csv file
-write.csv(employee_status_data, file = here::here("outputs/data/employee_status_data.csv"), row.names = TRUE)
-
-

### 2023-03-15 scripts/industry_sector.R

@@ -1,52 +1,0 @@
-#### Preamble ####
-# Purpose: data cleaning for industry sector
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-# loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-library(reshape2)
-
-# input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
-raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
-raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
-
-# extract the income and industry sector columns
-industry_sector_2018 <- raw_data_2018[, c("industry_sector", "income")]
-industry_sector_2016 <- raw_data_2016[, c("industry_sector", "income")]
-industry_sector_2014 <- raw_data_2014[, c("industry_sector", "income")]
-
-# combine three files into one
-industry_sector_data <- bind_rows(
-  industry_sector_2018 %>% mutate(year = "2018"),
-  industry_sector_2016 %>% mutate(year = "2016"),
-  industry_sector_2014 %>% mutate(year = "2014")
-)
-
-# only keep the income level 12
-# calcualte the number of each industry sector by year
-industry_sector_data <- industry_sector_data %>% 
-  filter(income == 12) %>% 
-  group_by(industry_sector, year) %>% 
-  summarize(people = n())
-
-# replace the industry sector code 
-industry_sector_replacements <- c("Government", "Private")
-industry_sector_data$industry_sector <- ifelse(industry_sector_data$industry_sector == 1, 
-                                               industry_sector_replacements[1],
-                                               industry_sector_replacements[2])
-
-# save the data frame as a csv file
-write.csv(industry_sector_data, file = here::here("outputs/data/industry_sector_data.csv"), row.names = TRUE)
-
-

### 2023-03-15 scripts/race.R

@@ -1,40 +1,0 @@
-#### Preamble ####
-# Purpose: data cleaning for race
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-# input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
-raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
-raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
-
-# extract the income and race columns
-race_2018 <- raw_data_2018[, c("race", "income")]
-race_2016 <- raw_data_2016[, c("race", "income")]
-race_2014 <- raw_data_2014[, c("race", "income")]
-
-# combine three files into one
-race_data <- bind_rows(
-  race_2018 %>% mutate(year = "2018"),
-  race_2016 %>% mutate(year = "2016"),
-  race_2014 %>% mutate(year = "2014")
-)
-
-# only keep the income level 12
-# calcualte the number of each race by year
-race_data <- race_data %>% 
-  filter(income == 12) %>% 
-  group_by(race, year) %>% 
-  summarize(people = n())
-
-# replace the college science course code 
-race_replacements <- c("White", "Black","Others")
-race_data$race <- ifelse(race_data$race == 1, race_replacements[1],
-                                ifelse(race_data$race == 2, race_replacements[2],
-                                       race_replacements[3]))
-
-# save the data frame as a csv file
-write.csv(race_data, file = here::here("outputs/data/race_data.csv"), row.names = TRUE)
-
-

### 2023-03-15 scripts/science_course.R

@@ -1,53 +1,0 @@
-#### Preamble ####
-# Purpose: data cleaning for college science course
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-
-# Loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-library(reshape2)
-
-# input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
-raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
-raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
-
-# extract the income and science course columns
-science_course_2018 <- raw_data_2018[, c("science_course", "income")]
-science_course_2016 <- raw_data_2016[, c("science_course", "income")]
-science_course_2014 <- raw_data_2014[, c("science_course", "income")]
-
-# combine three files into one
-science_course_data <- bind_rows(
-  science_course_2018 %>% mutate(year = "2018"),
-  science_course_2016 %>% mutate(year = "2016"),
-  science_course_2014 %>% mutate(year = "2014")
-)
-
-# only keep the income level 12
-# calcualte the number of college science course by year
-science_course_data <- science_course_data %>% 
-  filter(income == 12) %>% 
-  group_by(science_course, year) %>% 
-  summarize(people = n())
-
-# replace the college science course code 
-science_course_replacements <- c("Yes", "No")
-science_course_data$science_course <- ifelse(science_course_data$science_course == 1, 
-                                             science_course_replacements[1],
-                                             science_course_replacements[2])
-
-# save the data frame as a csv file
-write.csv(science_course_data, file = here::here("outputs/data/science_course_data.csv"), row.names = TRUE)
-
-

### 2023-03-15 scripts/sex.R

@@ -1,51 +1,0 @@
-#### Preamble ####
-# Purpose: data cleaning for sex
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-
-# Loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-library(reshape2)
-
-# input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
-raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
-raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
-
-# extract the income and sex columns
-sex_2018 <- raw_data_2018[, c("sex", "income")]
-sex_2016 <- raw_data_2016[, c("sex", "income")]
-sex_2014 <- raw_data_2014[, c("sex", "income")]
-
-# combine three files into one
-sex_data <- bind_rows(
-  sex_2018 %>% mutate(year = "2018"),
-  sex_2016 %>% mutate(year = "2016"),
-  sex_2014 %>% mutate(year = "2014")
-)
-
-# only keep the income level 12
-# calcualte the number of each sex by year
-sex_data <- sex_data %>% 
-  filter(income == 12) %>% 
-  group_by(sex, year) %>% 
-  summarize(people = n())
-
-# replace the sex code 
-sex_replacements <- c("Male", "Female")
-sex_data$sex <- ifelse(sex_data$sex == 1, 
-                       sex_replacements[1],
-                       sex_replacements[2])
-
-# save the data frame as a csv file
-write.csv(sex_data, file = here::here("outputs/data/sex_data.csv"), row.names = TRUE)

### 2023-03-15 scripts/test.R

@@ -1,40 +1,0 @@
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-library(reshape2)
-
-# Input data
-raw_data_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) 
-raw_data_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) 
-raw_data_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) 
-
-# extract the income and race columns
-employee_status_2018 <- raw_data_2018[, c("employee_status", "income")]
-employee_status_2016 <- raw_data_2016[, c("employee_status", "income")]
-employee_status_2014 <- raw_data_2014[, c("employee_status", "income")]
-
-# combine three files into one
-employee_status_data <- bind_rows(
-  employee_status_2018 %>% mutate(year = "2018"),
-  employee_status_2016 %>% mutate(year = "2016"),
-  employee_status_2014 %>% mutate(year = "2014")
-)
-
-# only keep the income level 12
-# calcualte the number of each emplyee status by year
-employee_status_data <- income_data %>% 
-  filter(income == 12) %>% 
-  group_by(employee_status, year) %>% 
-  summarize(freq = n())
-
-# replace the employee status code 
-employee_status_replacements <- c("Self-employed", "Someone else")
-employee_status_data$employee_status <- ifelse(employee_status_data$employee_status == 1, 
-                                                      employee_status_replacements[1],
-                                                      employee_status_replacements[2])

### 2023-03-15 outputs/paper/paper.qmd
final graph and data cleaning

@@
 
-```{r}
+# Introduction
+
+You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
+
+# Methodology
+
+```{r setup, include=FALSE}}
 knitr::opts_chunk$set(echo = FALSE)
@@
 ### Input data ###
-income_data_percentage<-read.csv(here::here("outputs/data/income_3years.csv"))
+happiness_data<-read.csv(here::here("outputs/data/happiness_data.csv"))
 sex_data<-read.csv(here::here("outputs/data/sex_data.csv"))
 race_data<-read.csv(here::here("outputs/data/race_data.csv"))
+degree_data<-read.csv(here::here("outputs/data/degree_data.csv"))
 employee_status_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
-industry_sector_data<-read.csv(here::here("outputs/data/industry_sector_data.csv"))
-science_course_data<-read.csv(here::here("outputs/data/science_course_data.csv"))
+marital_status_data<-read.csv(here::here("outputs/data/marital_status_data.csv"))
 
@@
 
-# Introduction
-
-You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
-
-```{r, fig.height=8}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-1
-#| fig-cap: "U.S. Income Distribution from 2014 to 2018"
-
-# Make sure the order of the income level is the order we wanted
-income_label <- c("LT $1000", "$1000 TO 2999", "$3000 TO 3999", "$4000 TO 4999", "$5000 TO 5999", "$6000 TO 6999", "$7000 TO 7999", "$8000 TO 9999", "$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 OR MORE")
-
-income_data_percentage$Income <- factor(income_data_percentage$Income, levels = income_label)
-
-# Define color palette with 3 colors
-fill_colors <- brewer.pal(n = 3, name = "RdYlBu")
-
-# Generate graph
-ggplot(income_data_percentage, aes(x = Income, y = Number, fill = as.factor(Year))) +
-  geom_bar(stat = "identity", position = "stack") +
-  scale_fill_manual(values = fill_colors) +
-  labs(x = "Income Level", y = "Total Number of People", title = "U.S. Income Distribution by Year") +
-  geom_text(aes(label = paste0(round(Percentage, 1), "%"), y = Number + 10), position = position_stack(vjust = 0.5), size = 2) +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8)) +
-  ylim(0, 2400)
-
-```
-
-# Data {#sec-data}
-
+# Results
 ```{r}
@@
 #| echo: false
-#| label: fig-2
-#| fig-cap: "U.S. Income Distribution by Sex"
-
-# calculate the percentage of each sex group for each year
-sex_data <- sex_data %>% 
-  group_by(year) %>% 
-  mutate(percentage = 100 * people / sum(people))
+#| label: fig-1Happiness
+#| fig-cap: "Changes in happiness levels from 1972 to 2021"
 
 # generate graph
-ggplot(sex_data, aes(x = year, y = people, 
-                                 fill = sex)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 1), size = 3) +
-  labs(x = "Year", y = "People", fill = "Sex",title = "sex")
+happy_colors <- c("#FF7F00", "#1F78B4", "#33A02C")
+
+ggplot(happiness_data, aes(x = Year, y = People, fill = Happy)) +
+  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
+  scale_fill_manual(values = happy_colors) +
+  scale_color_manual(values = happy_colors) +
+  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
+       title = "Changes in happiness levels from 1972 to 2021",
+       color = "Happy level")
 
@@
 #| echo: false
-#| label: fig-3
-#| fig-cap: "U.S. Income Distribution by Race"
+#| label: fig-2VeryHappyGender
+#| fig-cap: "Trends in happiness level (very happy) by gender from 1972 to 2021"
 
-# calculate the percentage of each race group for each year
-race_data <- race_data %>% 
-  group_by(year) %>% 
-  mutate(percentage = 100 * people / sum(people))
+# calculate the count of individuals for very happy level
+sex_data_happy1 <- sex_data %>% 
+  filter(Happy == "Very happy") %>% 
+  group_by(Sex, Year, Happy) %>% 
+  summarize(People = n())
 
 # generate graph
-ggplot(race_data, aes(x = year, y = people, 
-                                 fill = race)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 2), size = 3) +
-  labs(x = "Year", y = "People", fill = "Race",title = "race")
+sex_colors <- c("#E74C3C", "#2C3E50")
+
+ggplot(sex_data_happy1, aes(x = Year, y = People, fill = Sex)) +
+  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
+  scale_fill_manual(values = sex_colors) +
+  scale_color_manual(values = sex_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (very happy) by gender from 1972 to 2021")
+
+```
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-3PrettyHappyGender
+#| fig-cap: "Trends in happiness level (pretty happy) by gender from 1972 to 2021"
+
+# calculate the count of individuals for pretty happy level
+sex_data_happy2 <- sex_data %>% 
+  filter(Happy == "Pretty happy") %>% 
+  group_by(Sex, Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+sex_colors <- c("#E74C3C", "#2C3E50")
+
+ggplot(sex_data_happy2, aes(x = Year, y = People, fill = Sex)) +
+  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
+  scale_fill_manual(values = sex_colors) +
+  scale_color_manual(values = sex_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (pretty happy) by gender from 1972 to 2021")
 
@@
 #| echo: false
-#| label: fig-4
-#| fig-cap: "U.S. Income Distribution by Employee Status"
+#| label: fig-4NotTooHappyGender
+#| fig-cap: "Trends in happiness level (not too happy) by gender from 1972 to 2021"
 
-# calculate the percentage of each employee status group for each year
-employee_status_data <- employee_status_data %>% 
-  group_by(year) %>% 
-  mutate(percentage = 100 * people / sum(people))
+# calculate the count of individuals for not too happy level
+sex_data_happy3 <- sex_data %>% 
+  filter(Happy == "Not too happy") %>% 
+  group_by(Sex, Year, Happy) %>% 
+  summarize(People = n())
 
 # generate graph
-ggplot(employee_status_data, aes(x = year, y = people, 
-                                 fill = employee_status)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
-  labs(x = "Year", y = "People", fill = "Employee Status",title = "employee status")
+sex_colors <- c("#E74C3C", "#2C3E50")
+
+ggplot(sex_data_happy3, aes(x = Year, y = People, fill = Sex)) +
+  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
+  scale_fill_manual(values = sex_colors) +
+  scale_color_manual(values = sex_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (not too happy) by gender from 1972 to 2021")
 
@@
 #| echo: false
-#| label: fig-5
-#| fig-cap: "U.S. Income Distribution by Industry Sector"
+#| label: fig-5VeryHappyRace
+#| fig-cap: "Trends in happiness level (very happy) by race from 1972 to 2021"
 
-# calculate the percentage of each industry sector group for each year
-industry_sector_data <- industry_sector_data %>% 
-  group_by(year) %>% 
-  mutate(percentage = 100 * people / sum(people))
+# calculate the count of individuals for very happy level
+race_data_happy1 <- race_data %>% 
+  filter(Happy == "Very happy") %>% 
+  group_by(Race, Year, Happy) %>% 
+  summarize(People = n())
 
 # generate graph
-ggplot(industry_sector_data, aes(x = year, y = people, 
-                                 fill = industry_sector)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
-  labs(x = "Year", y = "People", fill = "Industry Sector",title = "industry sector")
+race_colors <- c("#3498DB", "#27AE60", "#F39C12")
+
+ggplot(race_data_happy1, aes(x = Year, y = People, fill = Race)) +
+  geom_line(aes(group = Race, color = Race), size = 1.2) +
+  scale_fill_manual(values = race_colors) +
+  scale_color_manual(values = race_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (very happy) by race from 1972 to 2021")
 
@@
 #| echo: false
-#| label: fig-6
-#| fig-cap: "U.S. Income Distribution by College Level Science Courses"
+#| label: fig-6PrettyHappyRace
+#| fig-cap: "Trends in happiness level (pretty happy) by race from 1972 to 2021"
 
-# calculate the percentage of taking college level science courses for each year
-science_course_data <- science_course_data %>% 
-  group_by(year) %>% 
-  mutate(percentage = 100 * people / sum(people))
+# calculate the count of individuals for pretty happy level
+race_data_happy2 <- race_data %>% 
+  filter(Happy == "Pretty happy") %>% 
+  group_by(Race, Year, Happy) %>% 
+  summarize(People = n())
 
 # generate graph
-ggplot(science_course_data, aes(x = year, y = people, 
-                                 fill = science_course)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  geom_text(aes(label = paste0(round(percentage), "%"), y = people + 5), position = position_dodge(width = 0.9), size = 3) +
-  labs(x = "Year", y = "People", fill = "Whether taking science course",title = "Science course")
+race_colors <- c("#3498DB", "#27AE60", "#F39C12")
+
+ggplot(race_data_happy2, aes(x = Year, y = People, fill = Race)) +
+  geom_line(aes(group = Race, color = Race), size = 1.2) +
+  scale_fill_manual(values = race_colors) +
+  scale_color_manual(values = race_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (pretty happy) by race from 1972 to 2021")
+
+```
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-7NotTooHappyRace
+#| fig-cap: "Trends in happiness level (not too happy) by race from 1972 to 2021"
+
+# calculate the count of individuals for not too happy level
+race_data_happy3 <- race_data %>% 
+  filter(Happy == "Not too happy") %>% 
+  group_by(Race, Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+race_colors <- c("#3498DB", "#27AE60", "#F39C12")
+
+ggplot(race_data_happy3, aes(x = Year, y = People, fill = Race)) +
+  geom_line(aes(group = Race, color = Race), size = 1.2) +
+  scale_fill_manual(values = race_colors) +
+  scale_color_manual(values = race_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (not too happy) by race from 1972 to 2021")
+
+```
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-8LessHighSchool
+#| fig-cap: "Changes in happiness levels of individuals with less than high school education from 1972 to 2021"
+
+# calcualte the people who have less than high school degree based on different level of happy
+degree_data_1 <- degree_data %>% 
+  filter(Degree == "Less than high school") %>% 
+  group_by(Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")
+
+ggplot(degree_data_1, aes(x = Year, y = People, fill = Happy)) +
+  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
+  scale_fill_manual(values = degree_colors) +
+  scale_color_manual(values = degree_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith less than high school education from 1972 to 2021", color = "Happy level")
+
+```
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-9HighSchool
+#| fig-cap: "Changes in happiness levels of individuals with high school education from 1972 to 2021"
+
+# calcualte the people who have high school degree based on different level of happy
+degree_data_2 <- degree_data %>% 
+  filter(Degree == "High school") %>% 
+  group_by(Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")
+
+ggplot(degree_data_2, aes(x = Year, y = People, fill = Happy)) +
+  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
+  scale_fill_manual(values = degree_colors) +
+  scale_color_manual(values = degree_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith high school education from 1972 to 2021", color = "Happy level")
+
+```
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-10AssociateCollege
+#| fig-cap: "Changes in happiness levels of individuals with associate/junior college education from 1972 to 2021"
+
+# calcualte the people who have associate/junior college degree based on different level of happy
+degree_data_3 <- degree_data %>% 
+  filter(Degree == "Associate/junior college") %>% 
+  group_by(Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")
+
+ggplot(degree_data_3, aes(x = Year, y = People, fill = Happy)) +
+  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
+  scale_fill_manual(values = degree_colors) +
+  scale_color_manual(values = degree_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith associate/junior college education from 1972 to 2021", color = "Happy level")
+```
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-11Bachelor
+#| fig-cap: "Changes in happiness levels of individuals with bachelor's education from 1972 to 2021"
+
+# calcualte the people who have Bachelor's degree based on different level of happy
+degree_data_4 <- degree_data %>% 
+  filter(Degree == "Bachelor's") %>% 
+  group_by(Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+degree_colors <- c("#E74C3C", "#F1C40F","#8E44AD")
+
+ggplot(degree_data_4, aes(x = Year, y = People, fill = Happy)) +
+  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
+  scale_fill_manual(values = degree_colors) +
+  scale_color_manual(values = degree_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith bachelor's education from 1972 to 2021", color = "Happy level")
+
+```
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-12Graduate
+#| fig-cap: "Changes in happiness levels of individuals with graduate education from 1972 to 2021"
+
+# calcualte the people who have graduate degree based on different level of happy
+degree_data_5 <- degree_data %>% 
+  filter(Degree == "Graduate") %>% 
+  group_by(Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+degree_colors <- c("#E74C3C", "#F1C40F","#8E44AD")
+
+ggplot(degree_data_5, aes(x = Year, y = People, fill = Happy)) +
+  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
+  scale_fill_manual(values = degree_colors) +
+  scale_color_manual(values = degree_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith graduate education from 1972 to 2021", color = "Happy level")
+
+```
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-13EmployeeStatusVeryHappy
+#| fig-cap: "Trends in happiness level (very happy) by employee status from 1972 to 2021"
+
+# calcualte the number of employee status for very happy
+employee_status_data_happy1 <- employee_status_data %>% 
+  filter(Happy == "Very happy") %>% 
+  group_by(Employee_status, Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+employee_status_colors <- c("#E74C3C", "#2C3E50")
+
+ggplot(employee_status_data_happy1, aes(x = Year, y = People, fill = Employee_status)) +
+  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
+  scale_fill_manual(values = employee_status_colors) +
+  scale_color_manual(values = employee_status_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (very happy) by employee status from 1972 to 2021", color = "Employee status")
+
+```
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-14EmployeeStatusPrettyHappy
+#| fig-cap: "Trends in happiness level (pretty happy) by employee status from 1972 to 2021"
+
+# calcualte the number of employee status for pretty happy
+employee_status_data_happy2 <- employee_status_data %>% 
+  filter(Happy == "Pretty happy") %>% 
+  group_by(Employee_status, Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+employee_status_colors <- c("#E74C3C", "#2C3E50")
+
+ggplot(employee_status_data_happy2, aes(x = Year, y = People, fill = Employee_status)) +
+  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
+  scale_fill_manual(values = employee_status_colors) +
+  scale_color_manual(values = employee_status_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (pretty happy) by employee status from 1972 to 2021", color = "Employee status")
+```
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-15EmployeeStatusNotTooHappy
+#| fig-cap: "Trends in happiness level (not too happy) by employee status from 1972 to 2021"
+
+# calcualte the number of employee status for not to happy
+employee_status_data_happy3 <- employee_status_data %>% 
+  filter(Happy == "Not too happy") %>% 
+  group_by(Employee_status, Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+employee_status_colors <- c("#E74C3C", "#2C3E50")
+
+ggplot(employee_status_data_happy3, aes(x = Year, y = People, fill = Employee_status)) +
+  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
+  scale_fill_manual(values = employee_status_colors) +
+  scale_color_manual(values = employee_status_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (not too happy) by employee status from 1972 to 2021", color = "Employee status")
+```
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-16MaritalStatusVeryHappy
+#| fig-cap: "Trends in Happiness Level (very happy) by Marital Status from 1972 to 2021"
+
+# calcualte the number of marital status for very happy
+marital_status_data_happy1 <- marital_status_data %>% 
+  filter(Happy == "Very happy") %>% 
+  group_by(Marital_status, Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")
+
+ggplot(marital_status_data_happy1, aes(x = Year, y = People, fill = Marital_status)) +
+  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
+  scale_fill_manual(values = marital_status_colors) +
+  scale_color_manual(values = marital_status_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (very happy) by marital status from 1972 to 2021", color = "Marital status")
+```
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-17MaritalStatusPrettyHappy
+#| fig-cap: "Trends in happiness level (pretty happy) by marital status from 1972 to 2021"
+
+# calcualte the number of marital status for pretty happy
+marital_status_data_happy2 <- marital_status_data %>% 
+  filter(Happy == "Pretty happy") %>% 
+  group_by(Marital_status, Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")
+
+ggplot(marital_status_data_happy2, aes(x = Year, y = People, fill = Marital_status)) +
+  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
+  scale_fill_manual(values = marital_status_colors) +
+  scale_color_manual(values = marital_status_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (pretty happy) by marital status from 1972 to 2021", color = "Marital status")
+```
+
+```{r}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-18MaritalStatusNotTooHappy
+#| fig-cap: "Trends in happiness level (not too happy) by marital status from 1972 to 2021"
+
+# calcualte the number of marital status for not too happy
+marital_status_data_happy3 <- marital_status_data %>% 
+  filter(Happy == "Not too happy") %>% 
+  group_by(Marital_status, Year, Happy) %>% 
+  summarize(People = n())
+
+# generate graph
+marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")
+
+ggplot(marital_status_data_happy3, aes(x = Year, y = People, fill = Marital_status)) +
+  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
+  scale_fill_manual(values = marital_status_colors) +
+  scale_color_manual(values = marital_status_colors) +
+  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (not too happy) by marital status from 1972 to 2021", color = "Marital status")
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
 

### 2023-03-15 scripts/00preparation.R
final graph and data cleaning

@@ -1,0 +1,52 @@
+#### Preamble ####
+# Purpose: Preparation for further data cleaning.
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+
+# Input raw data
+raw_data<-read_dta(here::here("inputs/data/gss7221_r3b.dta"))
+
+# Select only certain columns
+raw_data <-
+  data |>
+  select(
+    race,
+    year,
+    happy,
+    degree,
+    wrkslf,
+    marital,
+    sex
+  )
+
+# Delete the line if the value is NA
+raw_data <- na.omit(raw_data)
+
+# Rename the column
+raw_data <-
+  raw_data|>
+  rename(
+    Race = race,
+    Year = year,
+    Happy = happy,
+    Degree = degree,
+    Marital_status = marital,
+    Sex = sex,
+    Employee_status = wrkslf,
+  )
+
+# Save dat
+write.csv(raw_data, file = here::here("outputs/data/raw_data.csv"), row.names = FALSE)

### 2023-03-15 scripts/degree.R

@@ -1,0 +1,42 @@
+#### Preamble ####
+# Purpose: data cleaning for degree
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+
+# loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+## input data
+raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
+
+# extract the income and race columns
+degree_data <- raw_data[, c("Degree","Happy","Year")]
+
+# replace the degree code 
+degree_replacements <- c("Less than high school", "High school", "Associate/junior college", 
+                         "Bachelor's", "Graduate")
+degree_data$Degree <- ifelse(degree_data$Degree == 0, degree_replacements[1],
+                         ifelse(degree_data$Degree == 1, degree_replacements[2],
+                                ifelse(degree_data$Degree == 2, degree_replacements[3],
+                                       ifelse(degree_data$Degree == 3, degree_replacements[4],
+                                              ifelse(degree_data$Degree == 4, degree_replacements[5], NA)))))
+
+# replace the happy code
+happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
+degree_data$Happy <- ifelse(degree_data$Happy == 1, happy_replacements[1],
+                          ifelse(degree_data$Happy == 2, happy_replacements[2],
+                                 happy_replacements[3]))
+
+# save the data frame as a csv file
+write.csv(degree_data, file = here::here("outputs/data/degree_data.csv"), row.names = TRUE)

### 2023-03-15 scripts/employee_status.R

@@ -1,0 +1,41 @@
+#### Preamble ####
+# Purpose: data cleaning for employee status
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+# loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+## input data
+raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
+
+# extract the income and race columns
+employee_status_data <- raw_data[, c("Employee_status","Happy","Year")]
+
+# replace the employee status code 
+Employee_status_replacements <- c("Self-employed", "Someone else")
+employee_status_data$Employee_status <- ifelse(employee_status_data$Employee_status == 1, 
+                                               employee_status_replacements[1],
+                                               employee_status_replacements[2])
+
+# replace the happy code
+happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
+employee_status_data$Happy <- ifelse(employee_status_data$Happy == 1, happy_replacements[1],
+                          ifelse(employee_status_data$Happy == 2, happy_replacements[2],
+                                 happy_replacements[3]))
+
+# save the data frame as a csv file
+write.csv(employee_status_data, file = here::here("outputs/data/employee_status_data.csv"), row.names = TRUE)
+
+
+

### 2023-03-15 scripts/gdp.R

@@ -1,0 +1,77 @@
+#### Preamble ####
+# Purpose: data cleaning for three year income
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+# Input data
+income_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) %>% select(income)
+income_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) %>% select(income)
+income_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) %>% select(income)
+
+# Combine the data frames and add a column for the file name
+income_data <- bind_rows(
+  data.frame(year = "2018", income = income_2018),
+  data.frame(year = "2016", income = income_2016),
+  data.frame(year = "2014", income = income_2014))
+
+# group by year and income, count the number
+income_data_count <- income_data %>%
+  group_by(year, income) %>%
+  count() %>%
+  rename(number = n)
+
+# replace the income level code
+income_data_count <- income_data_count %>%
+  mutate(income = case_when(
+    income == 1 ~ "LT $1000",
+    income == 2 ~ "$1000 TO 2999",
+    income == 3 ~ "$3000 TO 3999",
+    income == 4 ~ "$4000 TO 4999",
+    income == 5 ~ "$5000 TO 5999",
+    income == 6 ~ "$6000 TO 6999",
+    income == 7 ~ "$7000 TO 7999",
+    income == 8 ~ "$8000 TO 9999",
+    income == 9 ~ "$10000 - 14999",
+    income == 10 ~ "$15000 - 19999",
+    income == 11 ~ "$20000 - 24999",
+    income == 12 ~ "$25000 OR MORE",
+    TRUE ~ as.character(income)
+  ))
+
+# rename the column
+income_data_count <-
+  income_data_count|>
+  rename(
+    Number = number,
+    Year = year,
+    Income = income
+  )
+
+# aggregate number column for each income level
+Number_Sum <- income_data_count %>%
+  group_by(Year) %>%
[…]

### 2023-03-15 scripts/happiness.R

@@ -1,0 +1,37 @@
+#### Preamble ####
+# Purpose: data cleaning for all year happiness
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+## input data
+raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
+
+# extract the income and race columns
+happiness_data <- raw_data[, c("Happy","Year")]
+
+# replace the happy code
+happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
+happiness_data$Happy <- ifelse(happiness_data$Happy == 1, happy_replacements[1],
+                         ifelse(happiness_data$Happy == 2, happy_replacements[2],
+                                happy_replacements[3]))
+
+# calcualte the number of people by year
+happiness_data <- happiness_data %>% 
+  group_by(Year, Happy) %>% 
+  summarize(People = n())
+
+# save the data frame as a csv file
+write.csv(happiness_data, file = here::here("outputs/data/happiness_data.csv"), row.names = TRUE)

### 2023-03-15 scripts/marital_status.R

@@ -1,0 +1,46 @@
+#### Preamble ####
+# Purpose: data cleaning for marital status
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+## input data
+raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
+
+# extract the income and race columns
+marital_status_data <- raw_data[, c("Marital_status","Happy","Year")]
+
+
+# replace the degree code 
+marital_status_replacements <- c("Married", "Widowed", "Divorced", 
+                                "Separated", "Never married")
+
+marital_status_data$Marital_status <- ifelse(marital_status_data$Marital_status == 1, marital_status_replacements[1],
+                                           ifelse(marital_status_data$Marital_status == 2, marital_status_replacements[2],
+                                                  ifelse(marital_status_data$Marital_status == 3, marital_status_replacements[3],
+                                                         ifelse(marital_status_data$Marital_status == 4, marital_status_replacements[4],
+                                                                ifelse(marital_status_data$Marital_status == 5, marital_status_replacements[5], 
+                                                                       NA)))))
+
+# replace the happy code
+happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
+marital_status_data$Happy <- ifelse(marital_status_data$Happy == 1, happy_replacements[1],
+                                   ifelse(marital_status_data$Happy == 2, happy_replacements[2],
+                                          happy_replacements[3]))
+
+# save the data frame as a csv file
+write.csv(marital_status_data, file = here::here("outputs/data/marital_status_data.csv"), row.names = TRUE)
+
+

### 2023-03-15 scripts/race.R

@@ -1,0 +1,38 @@
+#### Preamble ####
+# Purpose: data cleaning for race
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+# input data
+raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
+
+# extract the income and race columns
+race_data <- raw_data[, c("Race", "Happy","Year")]
+
+# replace the race code 
+race_replacements <- c("White", "Black","Others")
+race_data$Race <- ifelse(race_data$Race == 1, race_replacements[1],
+                         ifelse(race_data$Race == 2, race_replacements[2],
+                                race_replacements[3]))
+
+# replace the happy code
+happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
+race_data$Happy <- ifelse(race_data$Happy == 1, happy_replacements[1],
+                         ifelse(race_data$Happy == 2, happy_replacements[2],
+                                happy_replacements[3]))
+
+# save the data frame as a csv file
+write.csv(race_data, file = here::here("outputs/data/race_data.csv"), row.names = TRUE)

### 2023-03-15 scripts/sex.R

@@ -1,0 +1,39 @@
+#### Preamble ####
+# Purpose: data cleaning for sex
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+# Loading in relevant libraries
+library(tidyverse)
+library(dplyr, warn.conflicts = FALSE)
+library(tidyr)
+library(readxl)
+library(data.table)
+library(lubridate)
+library(haven)
+library(readxl)
+library(ggplot2)
+library(reshape2)
+
+## input data
+raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
+
+# extract the income and race columns
+sex_data <- raw_data[, c("Sex", "Happy","Year")]
+
+# replace the sex code 
+sex_replacements <- c("Male", "Female")
+sex_data$Sex <- ifelse(sex_data$Sex == 1, 
+                       sex_replacements[1],
+                       sex_replacements[2])
+
+# replace the happy code
+happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
+sex_data$Happy <- ifelse(sex_data$Happy == 1, happy_replacements[1],
+                          ifelse(sex_data$Happy == 2, happy_replacements[2],
+                                 happy_replacements[3]))
+
+# save the data frame as a csv file
+write.csv(sex_data, file = here::here("outputs/data/sex_data.csv"), row.names = TRUE)
+

### 2023-03-15 outputs/paper/paper.qmd
graph preview in pdf file

@@
 
-```{r setup, include=FALSE}}
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = FALSE)

### 2023-03-15 scripts/gdp.R

@@ -1,77 +1,0 @@
-#### Preamble ####
-# Purpose: data cleaning for three year income
-# Author: Rae Zhang, Faustine Fan, Myra Li
-# Data: 11 March 2023 
-# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
-
-
-# Loading in relevant libraries
-library(tidyverse)
-library(dplyr, warn.conflicts = FALSE)
-library(tidyr)
-library(readxl)
-library(data.table)
-library(lubridate)
-library(haven)
-library(readxl)
-library(ggplot2)
-library(reshape2)
-
-# Input data
-income_2018<-read.csv(here::here("outputs/data/preparation/GSS2018.csv")) %>% select(income)
-income_2016<-read.csv(here::here("outputs/data/preparation/GSS2016.csv")) %>% select(income)
-income_2014<-read.csv(here::here("outputs/data/preparation/GSS2014.csv")) %>% select(income)
-
-# Combine the data frames and add a column for the file name
-income_data <- bind_rows(
-  data.frame(year = "2018", income = income_2018),
-  data.frame(year = "2016", income = income_2016),
-  data.frame(year = "2014", income = income_2014))
-
-# group by year and income, count the number
-income_data_count <- income_data %>%
-  group_by(year, income) %>%
-  count() %>%
-  rename(number = n)
-
-# replace the income level code
-income_data_count <- income_data_count %>%
-  mutate(income = case_when(
-    income == 1 ~ "LT $1000",
-    income == 2 ~ "$1000 TO 2999",
-    income == 3 ~ "$3000 TO 3999",
-    income == 4 ~ "$4000 TO 4999",
-    income == 5 ~ "$5000 TO 5999",
-    income == 6 ~ "$6000 TO 6999",
-    income == 7 ~ "$7000 TO 7999",
-    income == 8 ~ "$8000 TO 9999",
-    income == 9 ~ "$10000 - 14999",
-    income == 10 ~ "$15000 - 19999",
-    income == 11 ~ "$20000 - 24999",
-    income == 12 ~ "$25000 OR MORE",
-    TRUE ~ as.character(income)
-  ))
-
-# rename the column
-income_data_count <-
-  income_data_count|>
-  rename(
-    Number = number,
-    Year = year,
-    Income = income
-  )
-
-# aggregate number column for each income level
-Number_Sum <- income_data_count %>%
-  group_by(Year) %>%
[…]

### 2023-03-15 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
readme file editted

--- 2023-03-15 outputs/paper/paper.qmd
@@
   labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
-       title = "Changes in happiness levels from 1972 to 2021",
+       title = "Fig1-Changes in happiness levels from 1972 to 2021",
        color = "Happy level")
@@
   scale_color_manual(values = sex_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (very happy) by gender from 1972 to 2021")
+  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Fig2-Trends in happiness level (very happy) by gender from 1972 to 2021")
 
@@
   scale_color_manual(values = sex_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (pretty happy) by gender from 1972 to 2021")
+  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Fig3-Trends in happiness level (pretty happy) by gender from 1972 to 2021")
 
@@
   scale_color_manual(values = sex_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (not too happy) by gender from 1972 to 2021")
+  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Fig4-Trends in happiness level (not too happy) by gender from 1972 to 2021")
 
@@
   scale_color_manual(values = race_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (very happy) by race from 1972 to 2021")
+  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Fig5-Trends in happiness level (very happy) by race from 1972 to 2021")
 
@@
   scale_color_manual(values = race_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (pretty happy) by race from 1972 to 2021")
+  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Fig6-Trends in happiness level (pretty happy) by race from 1972 to 2021")
 
@@
   scale_color_manual(values = race_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (not too happy) by race from 1972 to 2021")
+  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Fig7-Trends in happiness level (not too happy) by race from 1972 to 2021")
 
@@
   scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith less than high school education from 1972 to 2021", color = "Happy level")
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Fig8-Changes in happiness levels of individuals \nwith less than high school education from 1972 to 2021", color = "Happy level")
 
@@
   scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith high school education from 1972 to 2021", color = "Happy level")
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Fig9-Changes in happiness levels of individuals \nwith high school education from 1972 to 2021", color = "Happy level")
 
 ```
+
 ```{r}
@@
   scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith associate/junior college education from 1972 to 2021", color = "Happy level")
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Fig10-Changes in happiness levels of individuals \nwith associate/junior college education from 1972 to 2021", color = "Happy level")
 ```
@@
   scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith bachelor's education from 1972 to 2021", color = "Happy level")
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Fig11-Changes in happiness levels of individuals \nwith bachelor's education from 1972 to 2021", color = "Happy level")
 
@@
   scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith graduate education from 1972 to 2021", color = "Happy level")
+  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Fig12-Changes in happiness levels of individuals \nwith graduate education from 1972 to 2021", color = "Happy level")
 
@@
   scale_color_manual(values = employee_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (very happy) by employee status from 1972 to 2021", color = "Employee status")
+  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Fig13-Trends in happiness level (very happy) by employee status from 1972 to 2021", color = "Employee status")
 
@@
   scale_color_manual(values = employee_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (pretty happy) by employee status from 1972 to 2021", color = "Employee status")
+  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Fig14-Trends in happiness level (pretty happy) by employee status from 1972 to 2021", color = "Employee status")
 ```
@@
   scale_color_manual(values = employee_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (not too happy) by employee status from 1972 to 2021", color = "Employee status")
+  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Fig15-Trends in happiness level (not too happy) by employee status from 1972 to 2021", color = "Employee status")
 ```
@@
   scale_color_manual(values = marital_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (very happy) by marital status from 1972 to 2021", color = "Marital status")
+  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Fig16-Trends in happiness level (very happy) by marital status from 1972 to 2021", color = "Marital status")
 ```
@@
   scale_color_manual(values = marital_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (pretty happy) by marital status from 1972 to 2021", color = "Marital status")
+  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Fig17-Trends in happiness level (pretty happy) by marital status from 1972 to 2021", color = "Marital status")
 ```
@@
   scale_color_manual(values = marital_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (not too happy) by marital status from 1972 to 2021", color = "Marital status")
+  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Fig18-Trends in happiness level (not too happy) by marital status from 1972 to 2021", color = "Marital status")
 ```

### 2023-03-15 scripts/Fig13-15_employee_status_simulation.R
simulations

@@ -1,0 +1,20 @@
+#### Preamble ####
+# Purpose: Simulates the figure 13-15 raw data
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+library(janitor)
+library(tibble)
+library(tidyverse)
+
+set.seed(123)
+# Create vectors for each column
+year <- sample(1972:2021, 300, replace = TRUE)
+happy_level <- rep(1:3, length.out = 300)
+employee_status <- sample(c("Self-employed", "Someone else" ), 300, replace = TRUE)
+
+# Combine vectors into a tibble
+simulated_data_employee_status <- tibble(year, happy_level, employee_status)
+
+head(simulated_data_employee_status)

### 2023-03-15 scripts/Fig14-18_marital_status_simulation.R

@@ -1,0 +1,20 @@
+#### Preamble ####
+# Purpose: Simulates the figure 14-18 raw data
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+library(janitor)
+library(tibble)
+library(tidyverse)
+
+set.seed(123)
+# Create vectors for each column
+year <- sample(1972:2021, 300, replace = TRUE)
+happy_level <- rep(1:3, length.out = 300)
+marital_status <- sample(c("Married", "Widowed","Divorced","Separated","Never married" ), 300, replace = TRUE)
+
+# Combine vectors into a tibble
+simulated_data_marital_status <- tibble(year, happy_level, marital_status)
+
+head(simulated_data_marital_status)

### 2023-03-15 scripts/Fig1_happiness_simulation.R

@@ -1,0 +1,19 @@
+#### Preamble ####
+# Purpose: Simulates the figure 1 (Fig1-Changes in happiness levels from 1972 to 2021) raw data
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+library(janitor)
+library(tibble)
+library(tidyverse)
+
+# Create vectors for each column
+year <- rep(1972:2021, each = 3)
+happy_level <- rep(1:3, length(year)/3)
+people <- sample(1:100, length(year), replace = TRUE)
+
+# Combine vectors into a tibble
+simulated_data_happiness <- tibble(year, happy_level, people)
+
+head(simulated_data_happiness)

### 2023-03-15 scripts/Fig2-4_sex_simulation.R

@@ -1,0 +1,20 @@
+#### Preamble ####
+# Purpose: Simulates the figure 2-4 raw data
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+library(janitor)
+library(tibble)
+library(tidyverse)
+
+set.seed(123)
+# Create vectors for each column
+year <- sample(1972:2021, 300, replace = TRUE)
+happy_level <- rep(1:3, length.out = 300)
+sex <- sample(c("Female", "Male"), 300, replace = TRUE)
+
+# Combine vectors into a tibble
+simulated_data_sex <- tibble(year, happy_level, sex)
+
+head(simulated_data_sex)

### 2023-03-15 scripts/Fig5-7_race_simulation.R

@@ -1,0 +1,20 @@
+#### Preamble ####
+# Purpose: Simulates the figure 5-7 raw data
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+library(janitor)
+library(tibble)
+library(tidyverse)
+
+set.seed(123)
+# Create vectors for each column
+year <- sample(1972:2021, 300, replace = TRUE)
+happy_level <- rep(1:3, length.out = 300)
+race <- sample(c("White", "Black","Others"), 300, replace = TRUE)
+
+# Combine vectors into a tibble
+simulated_data_race <- tibble(year, happy_level, race)
+
+head(simulated_data_race)

### 2023-03-15 scripts/Fig8-12_degree_simulation.R

@@ -1,0 +1,21 @@
+#### Preamble ####
+# Purpose: Simulates the figure 8-12 raw data
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 11 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+library(janitor)
+library(tibble)
+library(tidyverse)
+
+set.seed(123)
+# Create vectors for each column
+year <- sample(1972:2021, 300, replace = TRUE)
+happy_level <- rep(1:3, length.out = 300)
+degree <- sample(c("Less than high school", "High school","Associate/junior college",
+                   "Bachelor's","Graduate"), 300, replace = TRUE)
+
+# Combine vectors into a tibble
+simulated_data_degree <- tibble(year, happy_level, degree)
+
+head(simulated_data_degree)

### 2023-03-15 scripts/test.R
test for cleaned data

@@ -1,0 +1,80 @@
+#### Preamble ####
+# Purpose: Test all cleaned data
+# Author: Rae Zhang, Faustine Fan, Myra Li
+# Data: 15 March 2023 
+# Contact: junwei.zhang@mail.utoronto.ca/yx.fan@mail.utoronto.ca/myra.li@mail.utoronto.ca
+
+library(tidyverse)
+
+########Input data ########
+figure1_data<-read.csv(here::here("outputs/data/happiness_data.csv"))
+figure2_4_data<-read.csv(here::here("outputs/data/sex_data.csv"))
+figure5_7_data<-read.csv(here::here("outputs/data/race_data.csv"))
+figure8_12_data<-read.csv(here::here("outputs/data/degree_data.csv"))
+figure13_15_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
+figure16_18_data<-read.csv(here::here("outputs/data/marital_status_data.csv"))
+
+########Figure 1########
+#test for year
+figure1_data$Year |> min() >= 1972
+figure1_data$Year |> max() <= 2021
+#test for happy level
+figure1_data$Happy |> 
+  unique() == c("Not too happy", "Pretty happy", "Very happy")
+#test for people
+figure1_data$People |> min() >= 0
+
+########Figure 2-4 ########
+#test for year
+figure2_4_data$Year |> min() >= 1972
+figure2_4_data$Year |> max() <= 2021
+#test for happy level
+figure2_4_data$Happy |> 
+  unique() == c("Not too happy", "Pretty happy", "Very happy")
+#test for sex
+figure2_4_data$Sex |> 
+  unique() == c("Female", "Male")
+
+########Figure 5-7 ########
+#test for year
+figure5_7_data$Year |> min() >= 1972
+figure5_7_data$Year |> max() <= 2021
+#test for happy level
+figure5_7_data$Happy |> 
+  unique() == c("Not too happy", "Pretty happy", "Very happy")
+#test for race
+figure5_7_data$Race |> 
+  unique() == c("White", "Black","Others")
+
+########Figure 8-12 ########
+#test for year
+figure8_12_data$Year |> min() >= 1972
+figure8_12_data$Year |> max() <= 2021
+#test for happy level
+figure8_12_data$Happy |> 
+  unique() == c("Not too happy", "Pretty happy", "Very happy")
+#test for race
+figure8_12_data$Degree |> 
[…]

### 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
data cleaning and graph

--- 2023-03-15 outputs/paper/paper.qmd
@@
 # loading in relevant libraries
-library(RColorBrewer)
+library(cowplot)
 library(tidyverse)
@@
 library(ggplot2)
+library(scales)
 
@@
 
-ggplot(happiness_data, aes(x = Year, y = People, fill = Happy)) +
+fig1 <- ggplot(happiness_data, aes(x = Year, y = People, fill = Happy)) +
   geom_line(aes(group = Happy, color = Happy), size = 1.2) +
@@
   labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
-       title = "Changes in happiness levels from 1972 to 2021",
        color = "Happy level")
+fig1 <- fig1  +
+  draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
+             fontface = "bold", 
+             size = 12,
+             x = 0.5,
+             y = 1.05)
+fig1
+```
+
+```{r, fig.height=6}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-2Gender
+#| fig-cap: "Trends in happiness levels for different genders from 1972 to 2021"
+
+# Only for female
+female_data <- filter(sex_data, Sex == "Female")
+gfemale <- ggplot(female_data, 
+       aes(y = Year, 
+           fill = Happy_level)) + 
+  geom_bar(position = "fill") + 
+  theme_classic() +
+  coord_flip()+
+  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
+       title = "Female",
+       color = "Happy level")
+
+# Only for male
+male_data <- filter(sex_data, Sex == "Male")
+gmale <- ggplot(male_data, 
+       aes(y = Year, 
+           fill = Happy_level)) + 
+  geom_bar(position = "fill") + 
+  theme_classic() +
+  coord_flip()+
+  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
+       title = "Male",
+       color = "Happy level")
+
+# group graphs together
+fig2 <- plot_grid(gfemale, gmale, ncol = 1,
+          label_size = 12,
+          label_x = 0.05,
+          label_y = 0.9,
+          label_fontface = "bold",
+          hjust = 0)
+
+fig2 <- fig2  +
+  draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
+             fontface = "bold", 
+             size = 16,
+             x = 0.5,
+             y = 1.05)
+fig2
+```
+
+```{r, fig.height=8}
+#| message: false
+#| warning: false
+#| echo: false
+#| label: fig-3Race
+#| fig-cap: "Trends in happiness levels for different races from 1972 to 2021"
+
+# Only for white
+white_data <- filter(race_data, Race == "White")
+gwhite <- ggplot(white_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "White people",
+           color = "Happy level")
+
+# Only for black
+black_data <- filter(race_data, Race == "Black")
+gblack <- ggplot(black_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip()+
+      labs(y = "Year", x = "Percentages", fill = "Happy Level", 
+           title = "black people",
+           color = "Happy level")
+
+# Only for others
+others_data <- filter(race_data, Race == "Others")
+gothers <- ggplot(others_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "other race",
+           color = "Happy level")
+
+# group graphs together
+fig3 <- plot_grid(gwhite, gblack, gothers, ncol = 1,
+          label_size = 14,
+          label_x = 0.05,
+          label_y = 0.9,
+          label_fontface = "bold",
+          hjust = 0)
+
+fig3 <- fig3  +
+  draw_label("Trends in happiness levels for different races from 1972 to 2021", 
+             fontface = "bold", 
+             size = 16,
+             x = 0.5,
+             y = 1.05)
+fig3
 
@@
 
-```{r}
+```{r, fig.height=16}
 #| message: false
@@
 #| echo: false
-#| label: fig-2VeryHappyGender
-#| fig-cap: "Trends in happiness level (very happy) by gender from 1972 to 2021"
+#| label: fig-4Degree
+#| fig-cap: "Trends in happiness levels for different degrees from 1972 to 2021"
 
-# calculate the count of individuals for very happy level
-sex_data_happy1 <- sex_data %>% 
-  filter(Happy == "Very happy") %>% 
-  group_by(Sex, Year, Happy) %>% 
-  summarize(People = n())
+# graph for less than high school
+less_high_school_data <- filter(degree_data, Degree == "Less than high school")
+glesshighschool <- ggplot(less_high_school_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Less than high school",
+           color = "Happy level")
 
-# generate graph
-sex_colors <- c("#E74C3C", "#2C3E50")
+# graph for high school
+high_school_data <- filter(degree_data, Degree == "High school")
+ghighschool <- ggplot(high_school_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "High school",
+           color = "Happy level")
 
-ggplot(sex_data_happy1, aes(x = Year, y = People, fill = Sex)) +
-  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
-  scale_fill_manual(values = sex_colors) +
-  scale_color_manual(values = sex_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (very happy) by gender from 1972 to 2021")
+# graph for associate/junior college
+junior_data <- filter(degree_data, Degree == "Associate/junior college")
+gjunior <- ggplot(junior_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Associate/junior college",
+           color = "Happy level")
 
+# graph for bachelor
+bachelor_data <- filter(degree_data, Degree == "Bachelor's")
+gbachelor <- ggplot(bachelor_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Bachelor",
+           color = "Happy level")
+
+# graph for graduate
+graduate_data <- filter(degree_data, Degree == "Graduate")
+graduate<- ggplot(graduate_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Graduate",
+           color = "Happy level")
+
+# group graphs together
+fig4 <- plot_grid(glesshighschool, ghighschool, gjunior, gbachelor, graduate, ncol = 1,
+          label_size = 14,
+          label_x = 0.05,
+          label_y = 0.9,
+          label_fontface = "bold",
+          hjust = 0)
+
+fig4 <- fig4  +
+  draw_label("Trends in happiness levels for different degrees from 1972 to 2021", 
+             fontface = "bold", 
+             size = 16,
+             x = 0.5,
+             y = 1.05)
+fig4
 ```
-```{r}
+
+```{r, fig.height=8}
 #| message: false
@@
 #| echo: false
-#| label: fig-3PrettyHappyGender
-#| fig-cap: "Trends in happiness level (pretty happy) by gender from 1972 to 2021"
+#| label: fig-5Employee
+#| fig-cap: "Trends in happiness levels for different emoloyee status from 1972 to 2021"
 
-# calculate the count of individuals for pretty happy level
-sex_data_happy2 <- sex_data %>% 
-  filter(Happy == "Pretty happy") %>% 
-  group_by(Sex, Year, Happy) %>% 
-  summarize(People = n())
+# graph for self-employee
+self_data <- filter(employee_status_data, Employee_status == "Self-employed")
+gself<- ggplot(self_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Self-employed",
+           color = "Happy level")
 
-# generate graph
-sex_colors <- c("#E74C3C", "#2C3E50")
+# graph for someone else
+someone_data <- filter(employee_status_data, Employee_status == "Someone else")
+gsomeone <- ggplot(someone_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Someone else",
+           color = "Happy level")
 
-ggplot(sex_data_happy2, aes(x = Year, y = People, fill = Sex)) +
-  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
-  scale_fill_manual(values = sex_colors) +
-  scale_color_manual(values = sex_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (pretty happy) by gender from 1972 to 2021")
+# group graphs together
+fig5 <- plot_grid(gself, gsomeone, ncol = 1,
+          label_size = 14,
+          label_x = 0.05,
+          label_y = 0.9,
+          label_fontface = "bold",
+          hjust = 0)
 
+fig5 <- fig5  +
+  draw_label("Trends in happiness levels for different emoloyee status from 1972 to 2021", 
+             fontface = "bold", 
+             size = 16,
+             x = 0.5,
+             y = 1.05)
+fig5
 ```
 
-```{r}
+```{r, fig.height=16}
 #| message: false
@@
 #| echo: false
-#| label: fig-4NotTooHappyGender
-#| fig-cap: "Trends in happiness level (not too happy) by gender from 1972 to 2021"
+#| label: fig-6Marital
+#| fig-cap: "Trends in happiness levels for different marital status from 1972 to 2021"
 
-# calculate the count of individuals for not too happy level
-sex_data_happy3 <- sex_data %>% 
-  filter(Happy == "Not too happy") %>% 
-  group_by(Sex, Year, Happy) %>% 
-  summarize(People = n())
+# graph for married
+married_data <- filter(marital_status_data, Marital_status == "Married")
+gmarried_data <- ggplot(married_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Married",
+           color = "Happy level")
 
-# generate graph
-sex_colors <- c("#E74C3C", "#2C3E50")
+# graph for widowed
+widowed_data <- filter(marital_status_data, Marital_status == "Widowed")
+gwidowed_data <- ggplot(widowed_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Widowed",
+           color = "Happy level")
 
-ggplot(sex_data_happy3, aes(x = Year, y = People, fill = Sex)) +
-  geom_line(aes(group = Sex, color = Sex), size = 1.2) +
-  scale_fill_manual(values = sex_colors) +
-  scale_color_manual(values = sex_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Sex", title = "Trends in happiness level (not too happy) by gender from 1972 to 2021")
+# graph for divorced
+divorced_data <- filter(marital_status_data, Marital_status == "Divorced")
+gdivorced_data <- ggplot(divorced_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Divorced",
+           color = "Happy level")
 
-```
+# graph for separated
+separated_data <- filter(marital_status_data, Marital_status == "Separated")
+gseparated_data <- ggplot(separated_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Separated",
+           color = "Happy level")
 
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-5VeryHappyRace
-#| fig-cap: "Trends in happiness level (very happy) by race from 1972 to 2021"
+# graph for never married
+never_married_data <- filter(marital_status_data, Marital_status == "Never married")
+gnever_married<- ggplot(never_married_data, 
+             aes(y = Year, 
+                 fill = Happy)) + 
+      geom_bar(position = "fill") + 
+      theme_classic() +
+      coord_flip() +
+      labs(y = "Year", x = "Percentage", fill = "Happy Level", 
+           title = "Never married",
+           color = "Happy level")
 
-# calculate the count of individuals for very happy level
-race_data_happy1 <- race_data %>% 
-  filter(Happy == "Very happy") %>% 
-  group_by(Race, Year, Happy) %>% 
-  summarize(People = n())
+# group graphs together
+fig6 <- plot_grid(gmarried_data, gnever_married, gwidowed_data, gdivorced_data, gseparated_data, ncol = 1,
+          label_size = 14,
+          label_x = 0.05,
+          label_y = 0.9,
+          label_fontface = "bold",
+          hjust = 0)
 
-# generate graph
-race_colors <- c("#3498DB", "#27AE60", "#F39C12")
-
-ggplot(race_data_happy1, aes(x = Year, y = People, fill = Race)) +
-  geom_line(aes(group = Race, color = Race), size = 1.2) +
-  scale_fill_manual(values = race_colors) +
-  scale_color_manual(values = race_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (very happy) by race from 1972 to 2021")
-
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-6PrettyHappyRace
-#| fig-cap: "Trends in happiness level (pretty happy) by race from 1972 to 2021"
-
-# calculate the count of individuals for pretty happy level
-race_data_happy2 <- race_data %>% 
-  filter(Happy == "Pretty happy") %>% 
-  group_by(Race, Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-race_colors <- c("#3498DB", "#27AE60", "#F39C12")
-
-ggplot(race_data_happy2, aes(x = Year, y = People, fill = Race)) +
-  geom_line(aes(group = Race, color = Race), size = 1.2) +
-  scale_fill_manual(values = race_colors) +
-  scale_color_manual(values = race_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (pretty happy) by race from 1972 to 2021")
-
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-7NotTooHappyRace
-#| fig-cap: "Trends in happiness level (not too happy) by race from 1972 to 2021"
-
-# calculate the count of individuals for not too happy level
-race_data_happy3 <- race_data %>% 
-  filter(Happy == "Not too happy") %>% 
-  group_by(Race, Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-race_colors <- c("#3498DB", "#27AE60", "#F39C12")
-
-ggplot(race_data_happy3, aes(x = Year, y = People, fill = Race)) +
-  geom_line(aes(group = Race, color = Race), size = 1.2) +
-  scale_fill_manual(values = race_colors) +
-  scale_color_manual(values = race_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Race", title = "Trends in happiness level (not too happy) by race from 1972 to 2021")
-
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-8LessHighSchool
-#| fig-cap: "Changes in happiness levels of individuals with less than high school education from 1972 to 2021"
-
-# calcualte the people who have less than high school degree based on different level of happy
-degree_data_1 <- degree_data %>% 
-  filter(Degree == "Less than high school") %>% 
-  group_by(Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")
-
-ggplot(degree_data_1, aes(x = Year, y = People, fill = Happy)) +
-  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
-  scale_fill_manual(values = degree_colors) +
-  scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith less than high school education from 1972 to 2021", color = "Happy level")
-
-```
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-9HighSchool
-#| fig-cap: "Changes in happiness levels of individuals with high school education from 1972 to 2021"
-
-# calcualte the people who have high school degree based on different level of happy
-degree_data_2 <- degree_data %>% 
-  filter(Degree == "High school") %>% 
-  group_by(Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")
-
-ggplot(degree_data_2, aes(x = Year, y = People, fill = Happy)) +
-  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
-  scale_fill_manual(values = degree_colors) +
-  scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith high school education from 1972 to 2021", color = "Happy level")
-
-```
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-10AssociateCollege
-#| fig-cap: "Changes in happiness levels of individuals with associate/junior college education from 1972 to 2021"
-
-# calcualte the people who have associate/junior college degree based on different level of happy
-degree_data_3 <- degree_data %>% 
-  filter(Degree == "Associate/junior college") %>% 
-  group_by(Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-degree_colors <- c("#E74C3C", "#F1C40F", "#8E44AD")
-
-ggplot(degree_data_3, aes(x = Year, y = People, fill = Happy)) +
-  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
-  scale_fill_manual(values = degree_colors) +
-  scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith associate/junior college education from 1972 to 2021", color = "Happy level")
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-11Bachelor
-#| fig-cap: "Changes in happiness levels of individuals with bachelor's education from 1972 to 2021"
-
-# calcualte the people who have Bachelor's degree based on different level of happy
-degree_data_4 <- degree_data %>% 
-  filter(Degree == "Bachelor's") %>% 
-  group_by(Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-degree_colors <- c("#E74C3C", "#F1C40F","#8E44AD")
-
-ggplot(degree_data_4, aes(x = Year, y = People, fill = Happy)) +
-  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
-  scale_fill_manual(values = degree_colors) +
-  scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith bachelor's education from 1972 to 2021", color = "Happy level")
-
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-12Graduate
-#| fig-cap: "Changes in happiness levels of individuals with graduate education from 1972 to 2021"
-
-# calcualte the people who have graduate degree based on different level of happy
-degree_data_5 <- degree_data %>% 
-  filter(Degree == "Graduate") %>% 
-  group_by(Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-degree_colors <- c("#E74C3C", "#F1C40F","#8E44AD")
-
-ggplot(degree_data_5, aes(x = Year, y = People, fill = Happy)) +
-  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
-  scale_fill_manual(values = degree_colors) +
-  scale_color_manual(values = degree_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Happy level", title = "Changes in happiness levels of individuals \nwith graduate education from 1972 to 2021", color = "Happy level")
-
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-13EmployeeStatusVeryHappy
-#| fig-cap: "Trends in happiness level (very happy) by employee status from 1972 to 2021"
-
-# calcualte the number of employee status for very happy
-employee_status_data_happy1 <- employee_status_data %>% 
-  filter(Happy == "Very happy") %>% 
-  group_by(Employee_status, Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-employee_status_colors <- c("#E74C3C", "#2C3E50")
-
-ggplot(employee_status_data_happy1, aes(x = Year, y = People, fill = Employee_status)) +
-  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
-  scale_fill_manual(values = employee_status_colors) +
-  scale_color_manual(values = employee_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (very happy) by employee status from 1972 to 2021", color = "Employee status")
-
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-14EmployeeStatusPrettyHappy
-#| fig-cap: "Trends in happiness level (pretty happy) by employee status from 1972 to 2021"
-
-# calcualte the number of employee status for pretty happy
-employee_status_data_happy2 <- employee_status_data %>% 
-  filter(Happy == "Pretty happy") %>% 
-  group_by(Employee_status, Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-employee_status_colors <- c("#E74C3C", "#2C3E50")
-
-ggplot(employee_status_data_happy2, aes(x = Year, y = People, fill = Employee_status)) +
-  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
-  scale_fill_manual(values = employee_status_colors) +
-  scale_color_manual(values = employee_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (pretty happy) by employee status from 1972 to 2021", color = "Employee status")
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-15EmployeeStatusNotTooHappy
-#| fig-cap: "Trends in happiness level (not too happy) by employee status from 1972 to 2021"
-
-# calcualte the number of employee status for not to happy
-employee_status_data_happy3 <- employee_status_data %>% 
-  filter(Happy == "Not too happy") %>% 
-  group_by(Employee_status, Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-employee_status_colors <- c("#E74C3C", "#2C3E50")
-
-ggplot(employee_status_data_happy3, aes(x = Year, y = People, fill = Employee_status)) +
-  geom_line(aes(group = Employee_status, color = Employee_status), size = 1.2) +
-  scale_fill_manual(values = employee_status_colors) +
-  scale_color_manual(values = employee_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Employee status", title = "Trends in happiness level (not too happy) by employee status from 1972 to 2021", color = "Employee status")
-```
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-16MaritalStatusVeryHappy
-#| fig-cap: "Trends in Happiness Level (very happy) by Marital Status from 1972 to 2021"
-
-# calcualte the number of marital status for very happy
-marital_status_data_happy1 <- marital_status_data %>% 
-  filter(Happy == "Very happy") %>% 
-  group_by(Marital_status, Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")
-
-ggplot(marital_status_data_happy1, aes(x = Year, y = People, fill = Marital_status)) +
-  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
-  scale_fill_manual(values = marital_status_colors) +
-  scale_color_manual(values = marital_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (very happy) by marital status from 1972 to 2021", color = "Marital status")
-```
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-17MaritalStatusPrettyHappy
-#| fig-cap: "Trends in happiness level (pretty happy) by marital status from 1972 to 2021"
-
-# calcualte the number of marital status for pretty happy
-marital_status_data_happy2 <- marital_status_data %>% 
-  filter(Happy == "Pretty happy") %>% 
-  group_by(Marital_status, Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")
-
-ggplot(marital_status_data_happy2, aes(x = Year, y = People, fill = Marital_status)) +
-  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
-  scale_fill_manual(values = marital_status_colors) +
-  scale_color_manual(values = marital_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (pretty happy) by marital status from 1972 to 2021", color = "Marital status")
-```
-
-```{r}
-#| message: false
-#| warning: false
-#| echo: false
-#| label: fig-18MaritalStatusNotTooHappy
-#| fig-cap: "Trends in happiness level (not too happy) by marital status from 1972 to 2021"
-
-# calcualte the number of marital status for not too happy
-marital_status_data_happy3 <- marital_status_data %>% 
-  filter(Happy == "Not too happy") %>% 
-  group_by(Marital_status, Year, Happy) %>% 
-  summarize(People = n())
-
-# generate graph
-marital_status_colors <- c("#E74C3C", "#F1C40F","#2ECC71","#3498DB","#8E44AD")
-
-ggplot(marital_status_data_happy3, aes(x = Year, y = People, fill = Marital_status)) +
-  geom_line(aes(group = Marital_status, color = Marital_status), size = 1.2) +
-  scale_fill_manual(values = marital_status_colors) +
-  scale_color_manual(values = marital_status_colors) +
-  labs(x = "Year", y = "Number of respondents", fill = "Marital status", title = "Trends in happiness level (not too happy) by marital status from 1972 to 2021", color = "Marital status")
-```
-
-```{r}
-
+fig6 <- fig6  +
+  draw_label("Trends in happiness levels for different emoloyee status from 1972 to 2021", 
+             fontface = "bold", 
+             size = 16,
+             x = 0.5,
+             y = 1.05)
+fig6
 ```

### 2023-03-16 scripts/degree.R

@@ -21,7 +21,7 @@ library(reshape2)
 raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
 
 # extract the income and race columns
-degree_data <- raw_data[, c("Degree","Happy","Year")]
+degree_data <- raw_data[, c("Degree","Happy_level","Year")]
 
 # replace the degree code 
 degree_replacements <- c("Less than high school", "High school", "Associate/junior college", 

### 2023-03-16 scripts/employee_status.R

@@ -20,13 +20,13 @@ library(reshape2)
 raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
 
 # extract the income and race columns
-employee_status_data <- raw_data[, c("Employee_status","Happy","Year")]
+employee_status_data <- raw_data[, c("Employee_status","Happy_level","Year")]
 
 # replace the employee status code 
 Employee_status_replacements <- c("Self-employed", "Someone else")
 employee_status_data$Employee_status <- ifelse(employee_status_data$Employee_status == 1, 
-                                               employee_status_replacements[1],
-                                               employee_status_replacements[2])
+                                               Employee_status_replacements[1],
+                                               Employee_status_replacements[2])
 
 # replace the happy code
 happy_replacements <- c("Very happy", "Pretty happy","Not too happy")

### 2023-03-16 scripts/marital_status.R

@@ -20,7 +20,7 @@ library(reshape2)
 raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
 
 # extract the income and race columns
-marital_status_data <- raw_data[, c("Marital_status","Happy","Year")]
+marital_status_data <- raw_data[, c("Marital_status","Happy_level","Year")]
 
 
 # replace the degree code 

### 2023-03-16 scripts/race.R

@@ -20,7 +20,7 @@ library(reshape2)
 raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
 
 # extract the income and race columns
-race_data <- raw_data[, c("Race", "Happy","Year")]
+race_data <- raw_data[, c("Race", "Happy_level","Year")]
 
 # replace the race code 
 race_replacements <- c("White", "Black","Others")

### 2023-03-16 scripts/sex.R

@@ -20,7 +20,7 @@ library(reshape2)
 raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
 
 # extract the income and race columns
-sex_data <- raw_data[, c("Sex", "Happy","Year")]
+sex_data <- raw_data[, c("Sex", "Happy_level","Year")]
 
 # replace the sex code 
 sex_replacements <- c("Male", "Female")
@@ -30,9 +30,11 @@ sex_data$Sex <- ifelse(sex_data$Sex == 1,
 
 # replace the happy code
 happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
-sex_data$Happy <- ifelse(sex_data$Happy == 1, happy_replacements[1],
-                          ifelse(sex_data$Happy == 2, happy_replacements[2],
+sex_data$Happy_level <- ifelse(sex_data$Happy_level == 1, happy_replacements[1],
+                          ifelse(sex_data$Happy_level == 2, happy_replacements[2],
                                  happy_replacements[3]))
+
+
 
 # save the data frame as a csv file
 write.csv(sex_data, file = here::here("outputs/data/sex_data.csv"), row.names = TRUE)

### 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
graph new changes

@@
 
-fig1 <- ggplot(happiness_data, aes(x = Year, y = People, fill = Happy)) +
-  geom_line(aes(group = Happy, color = Happy), size = 1.2) +
+fig1 <- ggplot(happiness_data, aes(x = Year, y = People, fill = Happy_level)) +
+  geom_line(aes(group = Happy_level, color = Happy_level), size = 1.2) +
   scale_fill_manual(values = happy_colors) +
@@
              aes(y = Year, 
-                 fill = Happy)) + 
+                 fill = Happy_level)) + 
       geom_bar(position = "fill") + 

### 2023-03-16 scripts/Fig1_happiness.R

@@ -20,17 +20,17 @@ library(reshape2)
 raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
 
 # extract the income and race columns
-happiness_data <- raw_data[, c("Happy","Year")]
+happiness_data <- raw_data[, c("Happy_level","Year")]
 
 # replace the happy code
 happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
-happiness_data$Happy <- ifelse(happiness_data$Happy == 1, happy_replacements[1],
-                         ifelse(happiness_data$Happy == 2, happy_replacements[2],
+happiness_data$Happy_level <- ifelse(happiness_data$Happy_level == 1, happy_replacements[1],
+                         ifelse(happiness_data$Happy_level == 2, happy_replacements[2],
                                 happy_replacements[3]))
 
 # calcualte the number of people by year
 happiness_data <- happiness_data %>% 
-  group_by(Year, Happy) %>% 
+  group_by(Year, Happy_level) %>% 
   summarize(People = n())
 
 # save the data frame as a csv file

### 2023-03-16 scripts/Fig3_race.R

@@ -30,8 +30,8 @@ race_data$Race <- ifelse(race_data$Race == 1, race_replacements[1],
 
 # replace the happy code
 happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
-race_data$Happy <- ifelse(race_data$Happy == 1, happy_replacements[1],
-                         ifelse(race_data$Happy == 2, happy_replacements[2],
+race_data$Happy_level <- ifelse(race_data$Happy_level == 1, happy_replacements[1],
+                         ifelse(race_data$Happy_level == 2, happy_replacements[2],
                                 happy_replacements[3]))
 
 # save the data frame as a csv file

### 2023-03-16 scripts/Fig4_degree.R

@@ -34,8 +34,8 @@ degree_data$Degree <- ifelse(degree_data$Degree == 0, degree_replacements[1],
 
 # replace the happy code
 happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
-degree_data$Happy <- ifelse(degree_data$Happy == 1, happy_replacements[1],
-                          ifelse(degree_data$Happy == 2, happy_replacements[2],
+degree_data$Happy_level <- ifelse(degree_data$Happy_level == 1, happy_replacements[1],
+                          ifelse(degree_data$Happy_level == 2, happy_replacements[2],
                                  happy_replacements[3]))
 
 # save the data frame as a csv file

### 2023-03-16 scripts/Fig5_employee_status.R

@@ -30,8 +30,9 @@ employee_status_data$Employee_status <- ifelse(employee_status_data$Employee_sta
 
 # replace the happy code
 happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
-employee_status_data$Happy <- ifelse(employee_status_data$Happy == 1, happy_replacements[1],
-                          ifelse(employee_status_data$Happy == 2, happy_replacements[2],
+employee_status_data$Happy_level <- ifelse(employee_status_data$Happy_level == 1, 
+                                           happy_replacements[1],
+                          ifelse(employee_status_data$Happy_level == 2, happy_replacements[2],
                                  happy_replacements[3]))
 
 # save the data frame as a csv file

### 2023-03-16 scripts/Fig6_marital_status.R

@@ -27,17 +27,23 @@ marital_status_data <- raw_data[, c("Marital_status","Happy_level","Year")]
 marital_status_replacements <- c("Married", "Widowed", "Divorced", 
                                 "Separated", "Never married")
 
-marital_status_data$Marital_status <- ifelse(marital_status_data$Marital_status == 1, marital_status_replacements[1],
-                                           ifelse(marital_status_data$Marital_status == 2, marital_status_replacements[2],
-                                                  ifelse(marital_status_data$Marital_status == 3, marital_status_replacements[3],
-                                                         ifelse(marital_status_data$Marital_status == 4, marital_status_replacements[4],
+marital_status_data$Marital_status <- ifelse(marital_status_data$Marital_status == 1, 
+                                             marital_status_replacements[1],
+                                           ifelse(marital_status_data$Marital_status == 2, 
+                                                  marital_status_replacements[2],
+                                                  ifelse(marital_status_data$Marital_status == 3, 
+                                                         marital_status_replacements[3],
+                                                         ifelse(marital_status_data$Marital_status == 4, 
+                                                                marital_status_replacements[4],
                                                                 ifelse(marital_status_data$Marital_status == 5, marital_status_replacements[5], 
                                                                        NA)))))
 
 # replace the happy code
 happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
-marital_status_data$Happy <- ifelse(marital_status_data$Happy == 1, happy_replacements[1],
-                                   ifelse(marital_status_data$Happy == 2, happy_replacements[2],
+marital_status_data$Happy_level <- ifelse(marital_status_data$Happy_level == 1, 
+                                    happy_replacements[1],
+                                   ifelse(marital_status_data$Happy_level == 2, 
+                                          happy_replacements[2],
                                           happy_replacements[3]))
 
 # save the data frame as a csv file

### 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
title, author changed for final paper

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Rae Zhang
+  - Faustine Fan
+  - Myra Li
+thanks: "Code and data are available at: https://github.com/JunweiZhang130/Happiness-in-America-A-Comprehensive-Analysis-of-the-Effects-of-Demographic-Factors.git."
 date: "`r Sys.time()`"
@@
 bibliography: references.bib
+
 ---

### 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
rename employee status to employee type

@@
              y = 1.05)
-fig1
 ```
@@
              y = 1.05)
-fig2
 ```
@@
              y = 1.05)
-fig3
-
 ```
@@
              y = 1.05)
-fig4
 ```
@@
              y = 1.05)
-fig5
 ```
@@
              y = 1.05)
-fig6
 ```

### 2023-03-16 scripts/00preparation.R

@@ -45,7 +45,7 @@ raw_data <-
     Degree = degree,
     Marital_status = marital,
     Sex = sex,
-    Employee_status = wrkslf,
+    Employee_type = wrkslf,
   )
 
 # Save dat

### 2023-03-16 scripts/Fig5_employee_type.R

@@ -20,23 +20,23 @@ library(reshape2)
 raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
 
 # extract the income and race columns
-employee_status_data <- raw_data[, c("Employee_status","Happy_level","Year")]
+employee_type_data <- raw_data[, c("Employee_status","Happy_level","Year")]
 
 # replace the employee status code 
-Employee_status_replacements <- c("Self-employed", "Someone else")
-employee_status_data$Employee_status <- ifelse(employee_status_data$Employee_status == 1, 
-                                               Employee_status_replacements[1],
-                                               Employee_status_replacements[2])
+Employee_type_replacements <- c("Self-employed", "Someone else")
+employee_type_data$Employee_type <- ifelse(employee_type_data$Employee_type == 1, 
+                                               Employee_type_replacements[1],
+                                               Employee_type_replacements[2])
 
 # replace the happy code
 happy_replacements <- c("Very happy", "Pretty happy","Not too happy")
-employee_status_data$Happy_level <- ifelse(employee_status_data$Happy_level == 1, 
+employee_type_data$Happy_level <- ifelse(employee_type_data$Happy_level == 1, 
                                            happy_replacements[1],
-                          ifelse(employee_status_data$Happy_level == 2, happy_replacements[2],
+                          ifelse(employee_type_data$Happy_level == 2, happy_replacements[2],
                                  happy_replacements[3]))
 
 # save the data frame as a csv file
-write.csv(employee_status_data, file = here::here("outputs/data/employee_status_data.csv"), row.names = TRUE)
+write.csv(employee_type_data, file = here::here("outputs/data/employee_status_data.csv"), row.names = TRUE)
 
 
 

### 2023-03-16 scripts/Fig5_employee_type_simulation.R

@@ -12,9 +12,9 @@ set.seed(123)
 # Create vectors for each column
 year <- sample(1972:2021, 300, replace = TRUE)
 happy_level <- rep(1:3, length.out = 300)
-employee_status <- sample(c("Self-employed", "Someone else" ), 300, replace = TRUE)
+employee_type <- sample(c("Self-employed", "Someone else" ), 300, replace = TRUE)
 
 # Combine vectors into a tibble
-simulated_data_employee_status <- tibble(year, happy_level, employee_status)
+simulated_data_employee_type <- tibble(year, happy_level, employee_type)
 
-head(simulated_data_employee_status)
+head(simulated_data_employee_type)

### 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
graph changed

@@
 ---
-title: "Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors"
+title: "Happiness in US "
+subtitle: "Analysis of the Effects of Demographic Factors during 1972 to 2021"
 author: 
@@
   - Myra Li
-thanks: "Code and data are available at: https://github.com/JunweiZhang130/Happiness-in-America-A-Comprehensive-Analysis-of-the-Effects-of-Demographic-Factors.git."
+thanks: "Code and data are available at: LINK."
 date: "`r Sys.time()`"
@@
 bibliography: references.bib
-
 ---
@@
 
-You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
+This analysis aims to gain a deeper understanding of the overall well-being and quality of life of people living in the United States by exploring the factors that contribute to happiness.
+While happiness is a subjective and personal concept, its impact on society is objective and significant.
+Identifying the factors that contribute to happiness is crucial for policymakers and individuals alike in making informed decisions that improve the quality of life for everyone.
+To achieve this, we focus on three main research questions:                 
+1) What are the trends of levels of happiness through the years 1972 to 2021?        
+2) What are the significant factors that affect the level of happiness people feel in society?                  
+3) How do these factors contribute to the trend of the levels of happiness through the years 1972 to 2021?              
+We hypothesize that factors such as education, social isolation, and work status will have an average impact on happiness levels, while demographic features such as race and gender will have a lesser impact.
+To provide readers with necessary background information, Section 2 outlines the survey data and collection methodology.
+In Section 3, the report presents the insights derived from analyzing the data.
+Section 4 provides a summary of the findings and outlines plans for further exploration of the dataset.
+Understanding the root causes of different levels of happiness can help policymakers and society develop effective strategies to improve citizens' quality of life.
+This analysis aims to provide valuable guidance for individuals to improve their satisfaction with life.
 
@@
 degree_data<-read.csv(here::here("outputs/data/degree_data.csv"))
-employee_status_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
+employee_type_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
 marital_status_data<-read.csv(here::here("outputs/data/marital_status_data.csv"))
@@
 #| label: fig-5Employee
-#| fig-cap: "Trends in happiness levels for different emoloyee status from 1972 to 2021"
+#| fig-cap: "Trends in happiness levels for different emoloyee type from 1972 to 2021"
 
 # graph for self-employee
-self_data <- filter(employee_status_data, Employee_status == "Self-employed")
+self_data <- filter(employee_type_data, Employee_type == "Self-employed")
 gself<- ggplot(self_data, 
@@
 # graph for someone else
-someone_data <- filter(employee_status_data, Employee_status == "Someone else")
+someone_data <- filter(employee_type_data, Employee_type == "Someone else")
 gsomeone <- ggplot(someone_data, 

### 2023-03-16 scripts/00preparation.R

@@ -20,8 +20,7 @@ library(ggplot2)
 raw_data<-read_dta(here::here("inputs/data/gss7221_r3b.dta"))
 
 # Select only certain columns
-raw_data <-
-  data |>
+raw_data <- raw_data %>%
   select(
     race,
     year,
@@ -41,7 +40,7 @@ raw_data <-
   rename(
     Race = race,
     Year = year,
-    Happy = happy,
+    Happy_level = happy,
     Degree = degree,
     Marital_status = marital,
     Sex = sex,

### 2023-03-16 scripts/Fig5_employee_type.R

@@ -20,7 +20,7 @@ library(reshape2)
 raw_data<-read.csv(here::here("outputs/data/raw_data.csv")) 
 
 # extract the income and race columns
-employee_type_data <- raw_data[, c("Employee_status","Happy_level","Year")]
+employee_type_data <- raw_data[, c("Employee_type","Happy_level","Year")]
 
 # replace the employee status code 
 Employee_type_replacements <- c("Self-employed", "Someone else")

### 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
add reference

@@
 Identifying the factors that contribute to happiness is crucial for policymakers and individuals alike in making informed decisions that improve the quality of life for everyone.
+
 To achieve this, we focus on three main research questions:
+
 1) What are the trends of levels of happiness through the years 1972 to 2021?
+
 2) What are the significant factors that affect the level of happiness people feel in society?
+
 3) How do these factors contribute to the trend of the levels of happiness through the years 1972 to 2021?
+
 We hypothesize that factors such as education, social isolation, and work status will have an average impact on happiness levels, while demographic features such as race and gender will have a lesser impact.
+
 To provide readers with necessary background information, Section 2 outlines the survey data and collection methodology.
@@
 
-# Methodology
+# Data 
+## Source and Methodology
+Our study is based on data collected from the US General Social Survey (GSS).
+The GSS is a nationally representative cross-sectional survey conducted since 1972, aimed at collecting data on contemporary American society to monitor and explain trends in opinions, attitudes, and behaviors.
+One of the unique features of the GSS is that it adapts questions from earlier surveys, enabling researchers to conduct comparisons for up to 80 years.
+Moreover, the GSS has taken great care to maintain survey methodology as comparable as possible over time, including keeping the same sampling approach and question wording.
+This approach helps to minimize potential changes due to variations in methodology and supports the study of trends in public opinion in the United States over time.
+
+The use of the GSS dataset in our research provides a robust and nationally representative sample of Americans, allowing us to examine the relationship between education and life satisfaction.
+This data source is widely recognized in the academic community for its comprehensive coverage of contemporary American society and its long-term approach to monitoring public opinion trends.
+
+To conduct this analysis, we utilized the R programming language and various packages for data cleaning, exploration, and visualization.
+The raw data was obtained as a Stata file and imported into R using the Haven package.
+The dataset included responses from 60,092 survey participants to 568 survey questions.
+To clean and analyze the data, we used the tidyverse and dplyr packages.
+To visualize the data, we utilized the ggplot2 package.
+
+The survey under consideration in this report is the General Social Survey, which has been conducted since 1972 by the NORC at the University of Chicago and funded by the National Science Foundation.
+The survey is designed to monitor public opinion and behavior in the United States and utilizes similar sampling and questioning approaches across its various rounds to ensure consistency.
+The GSS includes repeating modules and topical modules that may vary across rounds, and each participant is given a subset of these modules to respond to.
 
@@
 # Results
+
 ```{r}
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
add content

@@
 ---
-title: "Happiness in US "
+title: "Happiness in US"
 subtitle: "Analysis of the Effects of Demographic Factors during 1972 to 2021"
@@
 The GSS is a nationally representative cross-sectional survey conducted since 1972, aimed at collecting data on contemporary American society to monitor and explain trends in opinions, attitudes, and behaviors.
+
 One of the unique features of the GSS is that it adapts questions from earlier surveys, enabling researchers to conduct comparisons for up to 80 years.
@@
 
-The survey under consideration in this report is the General Social Survey, which has been conducted since 1972 by the NORC at the University of Chicago and funded by the National Science Foundation.
-The survey is designed to monitor public opinion and behavior in the United States and utilizes similar sampling and questioning approaches across its various rounds to ensure consistency.
-The GSS includes repeating modules and topical modules that may vary across rounds, and each participant is given a subset of these modules to respond to.
+## Survey Frame and Weakness
+Our survey has been carefully edited to improve upon existing questions from the original questionnaire and to add new questions that better capture the happiness of people in the USA.
+Specifically, we have addressed potential biases in questions and expanded the options to provide a more detailed exploration of the relationship between race and happiness.
+Additionally, we have added questions that focus on employment situations, income, COVID-19, mental health, future life aspirations, sense of fulfilment, and attitudes towards happiness.
 
+We believe that our survey will provide a more comprehensive and nuanced understanding of the factors that contribute to happiness in the USA.
+By improving the survey questions, we hope to provide policymakers and individuals with valuable insights into how to improve overall well-being and foster a happier society.
+
+## Strength and Weakness of Dataset 
+The strength of the GSS data is that the GSS uses a large and representative sample of the US population, which allows the broader population to contribute to the data set to eliminate the regional bias that might be caused in different provinces in the US.
+Meanwhile the GSS dat has been conducted consistently over a long period of time, allowing researchers to tracker changes especially in social attitudes.
+The GSS also uses standardized questions and methods, making it easier to compare data across different time periods and populations.
+
+The GSS data also has noticeable weakness.
+For example, while it covers a wide rangfe of topics, it may not capture all relevant factors for a particular research question.
+Meanwhile, with any survey, the GSS is subject to response bias, where certain individuals may be more or less likely to participate or answer certain questions.
+And the data only counts the US population, certain subgroups may be underrepresented or overrepresented in a sample.
+
+## Variables
+The 6 variables we selected from GSS for this research listed as follow:
+
+Race: the race respondent considered themselves are
+
+Happy: the happiness levels people felt at that moment
+
+Degree: the highest degree respondents had till that time
+
+Wrkslf: the employment status of the respondents
+
+Marital:  the marital status of respondents
+
+Sex: the sex respondents considered themselves are
+
+# Results
 ```{r setup, include=FALSE}
@@
 marital_status_data<-read.csv(here::here("outputs/data/marital_status_data.csv"))
-
 ```
 
-# Results
-
+## Overall Happiness
 ```{r}
@@
              y = 1.05)
+fig1
 ```
+Over the past five decades, @fig-1Happiness shows there has been an overall increase in happiness levels in the United States, though with significant fluctuations.
+Further analysis of this trend reveals that happiness is influenced by various factors, such as macroeconomic conditions, social changes, and personal circumstances.
+
+For instance, in the early 1990s, there was a considerable rise in happiness due to political and economic changes, such as the end of the Cold War, the emergence of neoliberalism, and the rise of a new Democratic Party (@ProtectingSoldiersandMothers).
+In contrast, the dot-com crash of the early 2000s led to a sharp decline in happiness levels that persisted until the economic recovery brought it back up to pre-crash levels.
+
+However, the global financial crisis and recession of 2007-2009 resulted in another decrease in happiness, as it impacted people's quality of life.
+More recently, the COVID-19 pandemic has had both positive and negative effects on happiness levels.
+While social isolation and economic uncertainty have decreased happiness, increased family and leisure time have had positive effects.
+
+We will examine how demographic factors such as sex, race, marital status, degree, and employment status influence happiness and why as follow.
+
+## Sex
 
@@
              y = 1.05)
+
+fig2
 ```
+In this analysis, we examined @fig-2Gender, which compares the proportion of different education levels for each sex group from 1972 to 2021.
+We observed that the distribution of overall happiness levels remains consistent over time, with "pretty happy" being the most common response for both male and female groups.
+
+A slight difference was noted in the mid-1970s, where the female group appeared slightly happier than the male group.
+However, the overall trends in happiness levels for both groups align with the trends shown in @fig-1Happiness.
+
+It is worth noting that our analysis is limited to the data presented in the figures, and further investigation may be required to understand any potential underlying factors contributing to these trends.
+
+## Race
+In @fig-3Race, we present the trends in happiness levels among different racial groups in America from 1972 to 2021.
+The figure shows that white people consistently report higher levels of happiness than other racial groups over the period, although this could be due to survivorship bias as the majority of respondents in the GSS dataset are white.
+It is important to note that all racial groups have become less happy in 2021, possibly due to the COVID-19 pandemic.
+
+Notably, the figure shows that black people have experienced drops in happiness levels in specific years, including 1980, 2008, and 2021.
+These drops in happiness levels may reflect the impact of significant events or social phenomena that occurred during those years and that disproportionately affected the happiness of the black race.
+Further research may be needed to explore the reasons for these trends and to determine the extent to which they are influenced by factors such as racism, economic inequality, and political events.
 
@@
              y = 1.05)
+
+fig3
 ```
+
+## Degree
+The trends in happiness levels across different education groups from 1972 to 2021 are presented in @fig-4Degree.
+It can be observed that individuals with a graduate and above education experience report the lowest levels of the "not too happy" response option and have the highest proportion of respondents selecting the "very happy" option across the years.
+Conversely, individuals with less than a high school education have the highest proportion of respondents selecting the "not too happy" response option among the different education groups over the years.
 
@@
              y = 1.05)
+
+fig4
 ```
+
+## Employee Type
+In general, there may not be significant differences between being self-employed or being employed by someone else in terms of overall trends and distribution.
+However, upon closer examination of @fig-5Employee, it becomes apparent that self-employed individuals tend to report higher levels of happiness than those who are employed by others.
+This could be attributed to several factors, such as the flexibility of their work schedule, the ability to reduce commute time, and fewer relationship issues.
+Overall, these benefits may contribute to a greater sense of control over one's work and personal life, which can positively impact overall happiness levels.
 
@@
              y = 1.05)
+
+fig5
 ```
 
+## Marital Status 
 ```{r, fig.height=16}
@@
              y = 1.05)
+
+fig6
 ```
+@fig-6Marital indicates that married individuals are more likely to report higher levels of happiness, with a greater proportion of respondents selecting "Pretty happy" and "Very happy" compared to those in other marital status categories.
+Interestingly, the proportion of respondents selecting "Very happy" in other groups is comparatively lower.
+Additionally, the distribution of happiness levels in the married group appears to be more stable and consistent over time compared to the other groups.
+
+It is worth noting that the COVID-19 pandemic had a significant impact on people's happiness levels, and married people’s level of happiness was more stable during this time.
+This may be attributed to the stability and security that marriage can provide during times of uncertainty.
 
@@
 
-# Results
+# Discussion
+## Race and Levels of Happiness 
+Upon analyzing @fig-2Gender, we observed that the overall distribution of happiness levels appears to be similar across different racial groups.
+However, there was a significant increase in the proportions of happiness levels reported by individuals identifying as Black or belonging to other racial groups during 1970 to 1990.
+This suggests that race identity may potentially influence people's happiness levels within a society.
+Individuals from diverse racial and ethnic groups may experience varying degrees of social and economic disadvantage, discrimination, and prejudice, which can impact their overall sense of well-being and happiness, leading to a major decrease or increase in self-reported happiness levels.
 
-# Discussion
+Research has shown that individuals who identify as racial minorities often experience more stress and lower levels of life satisfaction than those who identify as members of the dominant racial group (@FREDERICK20221).
+This can be due to factors such as experiences of racism, lack of representation in positions of power, and limited access to resources and opportunities.
+However, it is important to note that the impact of race identity on happiness is complex and multifaceted.
+The relationship between race and happiness can be influenced by factors such as social support, cultural identity, and personal values.
+Moreover, individual differences in personality, life circumstances, and coping strategies can also play a significant role in shaping the relationship between race identity and happiness.
+Therefore, while race identity can be a relevant factor in determining people's level of happiness, it is important to consider the context and individual differences in any analysis of this relationship.
+
+## Education levels and Happiness
+
+## Marital Status and level of Happiness
+
+## Weakness
+
+##Next Steps
 
@@
 
-Weaknesses and next steps should also be included.
-
 \newpage
@@
 
-# Appendix {.unnumbered}
+\newpage
+# Appendix
+A link to our survey exploring what will affect happiness in US can be found in the footnotes below.^[https://forms.gle/ZAoN1uFc9Gc24anh8]
 
+```{r, echo=FALSE, out.width="100%"}
+knitr::include_graphics(here::here("outputs/survey/survey_1.png"))
+knitr::include_graphics(here::here("outputs/survey/survey_2.png"))
+knitr::include_graphics(here::here("outputs/survey/survey_3.png"))
+knitr::include_graphics(here::here("outputs/survey/survey_4.png"))
+knitr::include_graphics(here::here("outputs/survey/survey_5.png"))
+knitr::include_graphics(here::here("outputs/survey/survey_6.png"))
+```
 # Additional details

### 2023-03-16 outputs/paper/Happiness-in-US-Analysis-of-the-Effects-of-Demographic-Factors-during-1972-to-2021.qmd
final paper version

--- 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
@@
   - Myra Li
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/JunweiZhang130/Happiness-in-US-Analysis-of-the-Effects-of-Demographic-Factors-during-1972-to-2021.git"
 date: "`r Sys.time()`"
 date-format: "D MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This essay explores the trends of happiness in the United States from 1972 to 2021 and identifies the significant factors that affect happiness levels.
+The report is based on the data gathered in US General Social Survey, and it investigates the impact of demographic features like sex, race, marital status, education, and employment types on happiness levels.
+The report hypothesizes that education, social isolation, and work status have a more significant impact on happiness levels than demographic features like race and gender.
+The COVID-19 pandemic has had overall negative effects on happiness levels.
+The report concludes that understanding the roots of different levels of happiness can help policymakers and society develop effective strategies to improve the quality of life for everyone."
 format: pdf
@@
   labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
-       color = "Happy level")
-fig1 <- fig1  +
+       color = "Happy level") +
   draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
@@
 ```
+
 Over the past five decades, @fig-1Happiness shows there has been an overall increase in happiness levels in the United States, though with significant fluctuations.
@@
 We will examine how demographic factors such as sex, race, marital status, degree, and employment status influence happiness and why as follow.
+\newpage
 
 ## Sex
-
 ```{r, fig.height=6}
@@
 ```
+
 In this analysis, we examined @fig-2Gender, which compares the proportion of different education levels for each sex group from 1972 to 2021.
@@
 ## Race
-In @fig-3Race, we present the trends in happiness levels among different racial groups in America from 1972 to 2021.
-The figure shows that white people consistently report higher levels of happiness than other racial groups over the period, although this could be due to survivorship bias as the majority of respondents in the GSS dataset are white.
-It is important to note that all racial groups have become less happy in 2021, possibly due to the COVID-19 pandemic.
-
-Notably, the figure shows that black people have experienced drops in happiness levels in specific years, including 1980, 2008, and 2021.
-These drops in happiness levels may reflect the impact of significant events or social phenomena that occurred during those years and that disproportionately affected the happiness of the black race.
-Further research may be needed to explore the reasons for these trends and to determine the extent to which they are influenced by factors such as racism, economic inequality, and political events.
-
 ```{r, fig.height=8}
@@
 ```
+In @fig-3Race (next page), we present the trends in happiness levels among different racial groups in America from 1972 to 2021.
+The figure shows that white people consistently report higher levels of happiness than other racial groups over the period, although this could be due to survivorship bias as the majority of respondents in the GSS dataset are white.
+It is important to note that all racial groups have become less happy in 2021, possibly due to the COVID-19 pandemic.
+
+Notably, the figure shows that black people have experienced drops in happiness levels in specific years, including 1980, 2008, and 2021.
+These drops in happiness levels may reflect the impact of significant events or social phenomena that occurred during those years and that disproportionately affected the happiness of the black race.
+Further research may be needed to explore the reasons for these trends and to determine the extent to which they are influenced by factors such as racism, economic inequality, and political events.
+\newpage
 
 ## Degree
-The trends in happiness levels across different education groups from 1972 to 2021 are presented in @fig-4Degree.
+The trends in happiness levels across different education groups from 1972 to 2021 are presented in @fig-4Degree (next page).
 It can be observed that individuals with a graduate and above education experience report the lowest levels of the "not too happy" response option and have the highest proportion of respondents selecting the "very happy" option across the years.
 Conversely, individuals with less than a high school education have the highest proportion of respondents selecting the "not too happy" response option among the different education groups over the years.
-
-```{r, fig.height=16}
+```{r, fig.height=12}
 #| message: false
@@
 ```
+\newpage
 
@@
 
-```{r, fig.height=8}
+```{r, fig.height=5}
 #| message: false
@@
 ```
+\newpage
 
 ## Marital Status 
+@fig-6Marital (next page) indicates that married individuals are more likely to report higher levels of happiness, with a greater proportion of respondents selecting "Pretty happy" and "Very happy" compared to those in other marital status categories.
+Interestingly, the proportion of respondents selecting "Very happy" in other groups is comparatively lower.
+Additionally, the distribution of happiness levels in the married group appears to be more stable and consistent over time compared to the other groups.
+
+It is worth noting that the COVID-19 pandemic had a significant impact on people's happiness levels, and married people’s level of happiness was more stable during this time.
+This may be attributed to the stability and security that marriage can provide during times of uncertainty.
+
 ```{r, fig.height=16}
@@
 ```
-@fig-6Marital indicates that married individuals are more likely to report higher levels of happiness, with a greater proportion of respondents selecting "Pretty happy" and "Very happy" compared to those in other marital status categories.
-Interestingly, the proportion of respondents selecting "Very happy" in other groups is comparatively lower.
-Additionally, the distribution of happiness levels in the married group appears to be more stable and consistent over time compared to the other groups.
-
-It is worth noting that the COVID-19 pandemic had a significant impact on people's happiness levels, and married people’s level of happiness was more stable during this time.
-This may be attributed to the stability and security that marriage can provide during times of uncertainty.
-
-```{r}
-
-```
-
-```{r}
-
-```
+\newpage
 
@@
 
-Research has shown that individuals who identify as racial minorities often experience more stress and lower levels of life satisfaction than those who identify as members of the dominant racial group (@FREDERICK20221).
+Research (@FREDERICK20221) has shown that individuals who identify as racial minorities often experience more stress and lower levels of life satisfaction than those who identify as members of the dominant racial group.
 This can be due to factors such as experiences of racism, lack of representation in positions of power, and limited access to resources and opportunities.
@@
 ## Education levels and Happiness
+Diener et al.
+(@Diener2013UniversalsAC) argues that education may foster a sense of personal growth, autonomy, and purpose, which can contribute to overall life satisfaction .
+This conclusion matches our observation of @fig-4Degree.
+The plot shows that higher levels of education can have a significant impact on people's level of happiness in the US.
+Education can provide individuals with important skills, knowledge, and opportunities that can improve their well-being and quality of life.
+For example, individuals with higher levels of education may have better job prospects, higher income, and greater access to resources and social support networks.
+These factors can contribute to higher levels of life satisfaction and happiness.
+One interesting fact we noticed is, less portion of people were negatively affected by the covid-19 in those who have above university degree.
+Which shows the skills gathered from higher education have advantages to maintain people's emotions.
+
+Education can provide individuals with important skills, knowledge, and opportunities that can improve their well-being and quality of life (@POWDTHAVEE201510).
+For example, individuals with higher levels of education tend to have better job prospects, higher income, and greater access to resources and social support networks.
+These factors can contribute to higher levels of life satisfaction and happiness.
+
+While our analysis suggests that education may have a positive impact on happiness levels, it is important to note that this relationship is not deterministic, and individual differences play a significant role in shaping one's happiness.
+As shown in @fig-3Race, even within the group of individuals with the highest level of education, there are still a significant number of respondents in every group selected "Not very happy" on the happiness scale.
+While education may contribute to higher levels of happiness, it is important to consider the complexity of this relationship and to recognize that happiness is a multifaceted construct that is influenced by a range of factors.
+This highlights the fact that education is not a guarantee of shaping one's sense of well-being.
 
 ## Marital Status and level of Happiness
+@fig-6Marital illustrates the relationship between marital status and happiness levels, showing that being married or in a committed relationship is associated with higher levels of life satisfaction and happiness compared to those who are single or divorced.
+
+There are several factors that may explain the relationship between marital status and happiness.
+Marital status plays a significant role in increasing happiness levels, as married people have more stable sexual partners (@MoneySexnHappiness).
+Marriage or a committed relationship can provide individuals with levels that may reflect the impact of significant events or social phenomena that social support, companionship, and a sense of belonging, which can contribute to feelings of happiness and well-being.
+Additionally, being in a stable and supportive relationship can provide individuals with a sense of security and stability that may contribute to their overall happiness.
 
 ## Weakness 
+One potential weakness of this paper is that the data used may not show all the responses to the survey row by row.
+Additionally, the dataset only represents responses from the United States, which limits its generalizability to countries with vastly different cultural backgrounds, such as those in Asia or South America.
 
-##Next Steps
+Furthermore, as the number of responses can vary from year to year, it's possible that the results may not be entirely objective, and errors in the dataset could occur.
+As a result, the conclusions drawn from this paper may not accurately reflect the overall factors that contribute to income inequality in society, and the findings may not be applicable globally.
+It's important to keep in mind that the conclusions drawn from this paper are based solely on data from the United States and may not be representative of the world as a whole.
+Therefore, caution should be exercised when applying these findings to other countries or regions.
 
-## First discussion point {#sec-first-point}
+## Next Steps
+When analyzing the relationship between marital status and happiness, it is important to consider the context and individual differences.
+Researchers should take into account factors such as the quality of the relationship, individual values and preferences, and other life circumstances that may impact one's level of happiness.
+By considering these factors, we can gain a more nuanced understanding of the relationship between marital status and happiness.
+Moving forward, there are several ways to expand on this study and enhance our understanding of income inequality.
+First, we encourage other researchers to replicate our study using the GSS data in the Github folder to test the validity and reliability of our findings.
+Additionally, we included a supplementary survey in appendix 1 that contains additional data we believe would be beneficial to this study.
+Collecting more data can enable us to conduct more in-depth research on the factors that contribute to income inequality using the GSS data.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-\newpage
-
-\appendix
+In addition, it would be valuable to compare our findings with those from other research papers that use national surveys in different countries to validate our conclusions and identify potential differences across regions.
+This could help us gain a more comprehensive understanding of the factors that contribute to income inequality on a global scale.
+By continuing to build on this research, we can work towards creating more equitable societies and improving the quality of life for individuals across the world.
 
@@
 A link to our survey exploring what will affect happiness in US can be found in the footnotes below.^[https://forms.gle/ZAoN1uFc9Gc24anh8]
+And you can access our survey through the QR code below.
+
+```{r, echo=FALSE, out.width="30%"}
+knitr::include_graphics(here::here("outputs/survey/QR_code.png"))
+```
 
@@
 ```
-# Additional details
-
-\newpage
-
-# References
+{{< pagebreak >}}
+## References

### 2023-03-16 outputs/paper/Happiness-in-US-Analysis-of-the-Effects-of-Demographic-Factors-during-1972-to-2021.qmd
changed title

@@
 ---
-title: "Happiness in US"
-subtitle: "Analysis of the Effects of Demographic Factors during 1972 to 2021"
+title: "Happiness in US: Analysis of the Effects of Demographic Factors during 1972 to 2021"
 author: 

### 2023-03-16 outputs/paper/Happiness-in-US-Analysis-of-the-Effects-of-Demographic-Factors-during-1972-to-2021.qmd
read me update

@@
 number-sections: true
+toc: true
 bibliography: references.bib
 ---
+\newpage
 
@@
 
-To conduct this analysis, we utilized the R programming language and various packages for data cleaning, exploration, and visualization.
-The raw data was obtained as a Stata file and imported into R using the Haven package.
+To conduct this analysis, we utilized the R programming language [@citeR] and various packages for data cleaning, exploration, and visualization.
+The raw data was obtained as a Stata file and imported into R using the haven [@haven] package.
 The dataset included responses from 60,092 survey participants to 568 survey questions.
-To clean and analyze the data, we used the tidyverse and dplyr packages.
-To visualize the data, we utilized the ggplot2 package.
+To clean and analyze the data, we used the tidyverse [@tidyverse], data table [@datatable] package, and dplyr [@dplyr] packages.
+To visualize the data, we utilized the ggplot2 [@ggplot2] package, scales [@scales] package.
+We used cowplot [@cowplot] package to group our graphs.
+And we used lubridate [@lubridate] package for auto-generated date.
 
@@
 library(tidyr)
-library(readxl)
 library(data.table)
@@
 library(haven)
-library(readxl)
 library(ggplot2)
@@
 
-For instance, in the early 1990s, there was a considerable rise in happiness due to political and economic changes, such as the end of the Cold War, the emergence of neoliberalism, and the rise of a new Democratic Party (@ProtectingSoldiersandMothers).
+For instance, in the early 1990s, there was a considerable rise in happiness due to political and economic changes, such as the end of the Cold War, the emergence of neoliberalism, and the rise of a new Democratic Party [@ProtectingSoldiersandMothers].
 In contrast, the dot-com crash of the early 2000s led to a sharp decline in happiness levels that persisted until the economic recovery brought it back up to pre-crash levels.
@@
 
-Research (@FREDERICK20221) has shown that individuals who identify as racial minorities often experience more stress and lower levels of life satisfaction than those who identify as members of the dominant racial group.
+Research [@FREDERICK20221] has shown that individuals who identify as racial minorities often experience more stress and lower levels of life satisfaction than those who identify as members of the dominant racial group.
 This can be due to factors such as experiences of racism, lack of representation in positions of power, and limited access to resources and opportunities.
@@
 Diener et al.
-(@Diener2013UniversalsAC) argues that education may foster a sense of personal growth, autonomy, and purpose, which can contribute to overall life satisfaction .
+[@Diener2013UniversalsAC] argues that education may foster a sense of personal growth, autonomy, and purpose, which can contribute to overall life satisfaction .
 This conclusion matches our observation of @fig-4Degree.
@@
 
-Education can provide individuals with important skills, knowledge, and opportunities that can improve their well-being and quality of life (@POWDTHAVEE201510).
+Education can provide individuals with important skills, knowledge, and opportunities that can improve their well-being and quality of life [@POWDTHAVEE201510].
 For example, individuals with higher levels of education tend to have better job prospects, higher income, and greater access to resources and social support networks.
@@
 There are several factors that may explain the relationship between marital status and happiness.
-Marital status plays a significant role in increasing happiness levels, as married people have more stable sexual partners (@MoneySexnHappiness).
+Marital status plays a significant role in increasing happiness levels, as married people have more stable sexual partners [@MoneySexnHappiness].
 Marriage or a committed relationship can provide individuals with levels that may reflect the impact of significant events or social phenomena that social support, companionship, and a sense of belonging, which can contribute to feelings of happiness and well-being.

### 2023-03-16 outputs/paper/Happiness and demographic factors.qmd
file name changed

--- 2023-03-16 outputs/paper/Happiness-in-US-Analysis-of-the-Effects-of-Demographic-Factors-during-1972-to-2021.qmd
@@
   coord_flip()+
-  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
+  labs(x = "Percentage", y = "Year", fill = "Happy Level", 
        title = "Female",
@@
   coord_flip()+
-  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
+  labs(x = "Percentage", y = "Year", fill = "Happy Level", 
        title = "Male",
@@
 \newpage
-# Appendix
+# Appendix {-}
 A link to our survey exploring what will affect happiness in US can be found in the footnotes below.^[https://forms.gle/ZAoN1uFc9Gc24anh8]
@@
 {{< pagebreak >}}
-## References
+# Reference {-}

### 2023-03-16 outputs/paper/Happiness and demographic factors.qmd
final version

@@
 ## Survey Frame and Weakness
-Our survey has been carefully edited to improve upon existing questions from the original questionnaire and to add new questions that better capture the happiness of people in the USA.
+Our survey has been carefully edited to improve upon existing questions from the original questionnaire and to add new questions that better capture the happiness of people in the US.
 Specifically, we have addressed potential biases in questions and expanded the options to provide a more detailed exploration of the relationship between race and happiness.
@@
 
-We believe that our survey will provide a more comprehensive and nuanced understanding of the factors that contribute to happiness in the USA.
+We believe that our survey will provide a more comprehensive and nuanced understanding of the factors that contribute to happiness in the US.
 By improving the survey questions, we hope to provide policymakers and individuals with valuable insights into how to improve overall well-being and foster a happier society.
